---
title: "Workshop Chapter 6: Goodness of fit"
author: "Lucas Hoogduin"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

<!-- GENERATED FILE: edit notebooks/goodness-of-fit/goodness-of-fit.Rmd in the private ada repository. -->


## Goodness of fit

### Exercise 6.1. Before we start

To run the exercises in this workshop, we need to load a few packages.

```{r setup}
library(ggplot2)         # Data visualization
# library(aicpa)           # Contains US SteamCo data
# library(car)             # Companion to Applied Regression
library(gridExtra)       # Arrange multiple grid-based plots
library(tidyr)           # Data tidying (reshape and transform)
library(tidyverse)
# library(corrplot)        # Correlation matrix visualization
library(lmtest)          # Diagnostic tests for linear models
# library(latex2exp)       # Use LaTeX expressions in plots
# library(scales)
library(kthemes)
theme_set(theme_grey())
```

### Exercise 6.2.  Customer satisfaction survey

The client satisfaction results from Table 6.3

```{r}
# Hypothesized population proportions from the base year
p <- c(0.40, 0.16, 0.44)
```

Results from the follow-up study

```{r}
# Observed frequencies from the follow-up study
observed <- c(15, 12, 13)

# Sample size
n <- sum(observed)
```

The expected frequencies are therefore obtained by multiplying these
proportions by the sample size.

```{r}
# Expected frequencies
(expected <- n * p)
```

The chi-square test statistic is

```{r}
# Chi-square test statistic (manual calculation)
(chi_sq <- sum((observed - expected)^2 / expected))
```

Calculation of the critical value

```{r}
# Degrees of freedom
(df <- length(observed) - 1)

# Critical value for alpha = 0.05
alpha <- 0.05
(critical_value <- qchisq(1 - alpha, df))
```

Calculation of the p-value

```{r}
(p_value <- pchisq(chi_sq, df, lower.tail = FALSE))
```

Using the built-in test

```{r}
chisq.test(x = observed, p = p)
```

### Benford's Law

```{r}
# Benford probabilities
digits <- 1:9
probabilities <- log10((digits + 1) / digits)

# Data frame
benford <- data.frame(
  Digit = factor(digits, levels = digits),
  Probability = probabilities
)

# Figure
ggplot(benford, aes(x = Digit, y = Probability)) +
  geom_col(fill = "#00338D", width = 0.7) +
  scale_y_continuous(
    limits = c(0, 0.35),
    breaks = seq(0, 0.35, by = 0.05),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    x = "First digit",
    y = "Probability"
  ) # +
  # theme_minimal(base_size = 12)
```

### Examples from Benford

```{r}
library(tidyverse)

df <- tribble(
  ~Digit, ~Rivers, ~AmLeague, ~CostData, ~ReadersDigest, ~MolWgt, ~Average,
  1, 31.0, 32.7, 32.4, 33.4, 26.7, 31.2,
  2, 16.4, 17.6, 18.8, 18.5, 25.2, 19.3,
  3, 10.7, 12.6, 10.1, 12.4, 15.4, 12.2,
  4, 11.3,  9.8, 10.1,  7.5, 10.8,  9.9,
  5,  7.2,  7.4,  9.8,  7.1,  6.7,  7.6,
  6,  8.6,  6.4,  5.5,  6.5,  5.1,  6.4,
  7,  5.5,  4.9,  4.7,  5.5,  4.1,  4.9,
  8,  4.2,  5.6,  5.5,  4.9,  2.8,  4.6,
  9,  5.1,  3.0,  3.1,  4.2,  3.2,  3.7
)

df_long <- pivot_longer(
  df,
  -Digit,
  names_to = "Series",
  values_to = "Percentage"
)

ggplot(df_long,
       aes(x = Digit,
           y = Percentage,
           color = Series)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = 1:9) +
  labs(
    x = "First digit",
    y = "Percentage"
  ) 
```

### Comparison of average with theoretical

```{r}
# Benford probabilities (%)
benford <- 100 * probabilities

# Average from the table
average <- c(31.2, 19.3, 12.2, 9.9, 7.6, 6.4, 4.9, 4.6, 3.7)

# Data frame
df <- data.frame(
  Digit = rep(digits, 2),
  Percentage = c(benford, average),
  Series = rep(c("Benford", "Average"), each = 9)
)

# Plot
ggplot(df,
       aes(x = Digit,
           y = Percentage,
           color = Series,
           linetype = Series)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = 1:9) +
  labs(
    x = "First digit",
    y = "Percentage"
  ) 
```
## Inventory data

```{r}
# Benford probabilities (%)
inv_expected <- 300 * probabilities

# Observed frequencies inventories
inv_observed <- c(86, 48, 23, 32, 24, 36, 19, 18, 14)

# Data frame
df_inv <- data.frame(
  Digit = rep(digits, 2),
  Frequency = c(inv_expected, inv_observed),
  Series = rep(c("Expected", "Observed"), each = 9)
)

# Plot
inv_plot <- ggplot(df_inv,
       aes(x = Digit,
           y = Frequency,
           color = Series,
           linetype = Series)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = 1:9) +
  theme(legend.position = "top") +
  guides(color = guide_legend(title = NULL), linetype = guide_legend(title = NULL)) + # Remove legend titles
  labs(
    x = "First digit",
    y = "Frequency"
  ) 

inv_plot
```
Calculation of the test statistic

```{r}
# Expected
round(inv_expected, 2)
```
```{r}
# Obs - Exp squared
round((inv_observed - inv_expected)^2, 2)
```
```{r}
# Obs - Exp squared divided by expected
round(((inv_observed - inv_expected)^2 / inv_expected), 2)
sum(round(((inv_observed - inv_expected)^2 / inv_expected), 2))
```

Critical region
```{r}
# Parameters
df <- 8  # Degrees of freedom
x <- seq(0, 20, by = 0.1)  # Range of x values
tail_start <- 15.51  # Starting point for the shaded tail

# Create a data frame for the chi-square distribution
data <- data.frame(
  x = x,
  y = dchisq(x, df = df)
)

# Create the plot
ggplot(data, aes(x = x, y = y)) +
  geom_line(color = "#00338D", size = 1) +
  geom_area(
    data = subset(data, x >= tail_start),
    aes(x = x, y = y),
    fill = "#00338D",
    alpha = 1
  ) +
  labs(x="x", y="density")
```


Using the built-in test

```{r}
chisq.test(x = inv_observed, p = probabilities)
```
## Discrete Uniform Distribution

```{r}
# First digit probabilities
first_digit <- data.frame(
  digit = 1:9,
  probability = log10(1 + 1/1:9),
  position = "First"
)

# Second digit probabilities
second_digit <- data.frame(
  digit = 0:9,
  probability = sapply(0:9, function(d) sum(log10(1 + 1/(10 * 1:9 + d)))),
  position = "Second"
)

# Third digit probabilities
third_digit <- data.frame(
  digit = 0:9,
  probability = sapply(0:9, function(d) sum(log10(1 + 1/(10 * 10:99 + d)))),
  position = "Third"
)

# Combine data
benford_data <- bind_rows(first_digit, second_digit, third_digit)

# Create line plots
ggplot(benford_data, aes(x = digit, y = probability, group = position)) +
  geom_line() +
  geom_point() +
  facet_wrap(~position, ncol = 1, scales = "free_x") +
  labs(
    x = "Digit",
    y = "Probability"
  ) +
  theme(legend.position = "none")
```




```{r}
library(ggplot2)
library(grid)
library(gridExtra)

nres <- ussteamco.mod.3$residuals / sd(ussteamco.mod.3$residuals)
model3 <- as.data.frame(nres)

# Function to create a histogram with a given number of bins
create_hist <- function(bins) {
  ggplot(model3, aes(nres)) +
    geom_histogram(aes(y = after_stat(density)),
                   breaks = seq(min(nres), max(nres), length.out = bins + 1),
                   fill = "#00338D", color = "black") +
    stat_function(fun = dnorm,
                  args = list(mean = mean(model3$nres),
                              sd = sd(model3$nres)),
                  color = "red", size = 1) +
    labs(title = paste("Histogram with", bins, "bins"),
         x = "Standardized Residuals", y = "Density") +
    theme_minimal()
}

# Create the four histograms
hist_6 <- create_hist(6)
hist_7 <- create_hist(7)
hist_8 <- create_hist(8)
hist_9 <- create_hist(9)

# Arrange in a 2x2 grid using grid
grid.arrange(hist_6, hist_7, hist_8, hist_9, ncol = 2)
```
