---
title: "Workshop Chapter 6: Goodness of fit"
author: "Lucas Hoogduin"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

<!-- GENERATED FILE: edit notebooks/support/goodness-of-fit/support.Rmd in the private ada repository. -->

**Volume 1 workbooks v0.1.0, published 2026-08-19.**


## Goodness of fit

### Exercise 6.1. Before we start

To run the exercises in this workshop, we need to load a few packages.

```{r setup}
library(ggplot2) # Data visualization
# library(aicpa)           # Contains US SteamCo data
# library(car)             # Companion to Applied Regression
library(gridExtra) # Arrange multiple grid-based plots
library(tidyr) # Data tidying (reshape and transform)
library(tidyverse)
# library(corrplot)        # Correlation matrix visualization
library(lmtest) # Diagnostic tests for linear models
# library(latex2exp)       # Use LaTeX expressions in plots
# library(scales)
ada_colour <- function(name) {
  palette <- c(
    blue = "#3c4b99",
    mediumblue = "#6098d2",
    lightblue = "#62b8e7",
    red = "#c93f55",
    yellow = "#eacc62",
    green = "#469d76",
    purple = "#924099",
    lightpurple = "#df9ed4",
    blue50 = "#d3d8ef",
    lightpink = "#e36877"
  )
  unname(palette[[name]])
}
theme_set(ggplot2::theme_bw())
```

### Exercise 6.2.  Customer satisfaction survey

The client satisfaction results from the manuscript client-satisfaction table

```{r}
# Hypothesized population proportions from the base year
p <- c(0.40, 0.16, 0.44)
```

Results from the follow-up study

```{r}
# Observed frequencies from the follow-up study
observed <- c(15, 12, 13)

# Sample size
n <- sum(observed)
```

The expected frequencies are therefore obtained by multiplying these
proportions by the sample size.

```{r}
# Expected frequencies
(expected <- n * p)
```

The chi-square test statistic is

```{r}
# Chi-square test statistic (manual calculation)
(chi_sq <- sum((observed - expected)^2 / expected))
```

Calculation of the critical value

```{r}
# Degrees of freedom
(df <- length(observed) - 1)

# Critical value for alpha = 0.05
alpha <- 0.05
(critical_value <- qchisq(1 - alpha, df))
```

Calculation of the p-value

```{r}
(p_value <- pchisq(chi_sq, df, lower.tail = FALSE))
```

Using the built-in test

```{r}
chisq.test(x = observed, p = p)
```

```{r}
# Benford probabilities
digits <- 1:9
probabilities <- log10((digits + 1) / digits)

# Data frame
benford <- data.frame(
  Digit = factor(digits, levels = digits),
  Probability = probabilities
)

# Figure
benford_plot <- ggplot(benford, aes(x = Digit, y = Probability)) +
  geom_col(fill = ada_colour("blue"), width = 0.7) +
  scale_y_continuous(
    limits = c(0, 0.35),
    breaks = seq(0, 0.35, by = 0.05),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    x = "First digit",
    y = "Probability"
  ) # +
# ggplot2::theme_bw() is set globally for ADA figures

benford_plot
```

```{r}
library(tidyverse)

df <- tribble(
  ~Digit, ~Rivers, ~AmLeague, ~CostData, ~ReadersDigest, ~MolWgt, ~Average,
  1, 31.0, 32.7, 32.4, 33.4, 26.7, 31.2,
  2, 16.4, 17.6, 18.8, 18.5, 25.2, 19.3,
  3, 10.7, 12.6, 10.1, 12.4, 15.4, 12.2,
  4, 11.3, 9.8, 10.1, 7.5, 10.8, 9.9,
  5, 7.2, 7.4, 9.8, 7.1, 6.7, 7.6,
  6, 8.6, 6.4, 5.5, 6.5, 5.1, 6.4,
  7, 5.5, 4.9, 4.7, 5.5, 4.1, 4.9,
  8, 4.2, 5.6, 5.5, 4.9, 2.8, 4.6,
  9, 5.1, 3.0, 3.1, 4.2, 3.2, 3.7
)

df_long <- pivot_longer(
  df,
  -Digit,
  names_to = "Series",
  values_to = "Percentage"
)

benford_examples_plot <- ggplot(
  df_long,
  aes(
    x = Digit,
    y = Percentage,
    color = Series
  )
) +
  ggplot2::theme_bw() +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c(
    "Rivers" = ada_colour("blue"),
    "AmLeague" = ada_colour("red"),
    "CostData" = ada_colour("yellow"),
    "ReadersDigest" = ada_colour("green"),
    "MolWgt" = ada_colour("purple"),
    "Average" = ada_colour("lightpurple")
  )) +
  scale_x_continuous(breaks = 1:9) +
  labs(
    x = "First digit",
    y = "Percentage"
  )

benford_examples_plot
```

```{r}
# Benford probabilities (%)
benford <- 100 * probabilities

# Average from the table
average <- c(31.2, 19.3, 12.2, 9.9, 7.6, 6.4, 4.9, 4.6, 3.7)

# Data frame
df <- data.frame(
  Digit = rep(digits, 2),
  Percentage = c(benford, average),
  Series = rep(c("Benford", "Average"), each = 9)
)

# Plot
benford_average_plot <- ggplot(
  df,
  aes(
    x = Digit,
    y = Percentage,
    color = Series,
    linetype = Series
  )
) +
  ggplot2::theme_bw() +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c(
    "Benford" = ada_colour("blue"),
    "Average" = ada_colour("red")
  )) +
  scale_x_continuous(breaks = 1:9) +
  labs(
    x = "First digit",
    y = "Percentage"
  )

benford_average_plot
```
```{r}
# Benford probabilities (%)
inv_expected <- 300 * probabilities

# Observed frequencies inventories
inv_observed <- c(86, 48, 23, 32, 24, 36, 19, 18, 14)

# Data frame
df_inv <- data.frame(
  Digit = rep(digits, 2),
  Frequency = c(inv_expected, inv_observed),
  Series = rep(c("Expected", "Observed"), each = 9)
)

# Plot
inv_plot <- ggplot(
  df_inv,
  aes(
    x = Digit,
    y = Frequency,
    color = Series,
    linetype = Series
  )
) +
  ggplot2::theme_bw() +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c(
    "Expected" = ada_colour("blue"),
    "Observed" = ada_colour("red")
  )) +
  scale_x_continuous(breaks = 1:9) +
  theme(legend.position = "top") +
  guides(color = guide_legend(title = NULL), linetype = guide_legend(title = NULL)) + # Remove legend titles
  labs(
    x = "First digit",
    y = "Frequency"
  )

inv_plot
```
Calculation of the test statistic

```{r}
# Expected
round(inv_expected, 2)
```
```{r}
# Obs - Exp squared
round((inv_observed - inv_expected)^2, 2)
```
```{r}
# Obs - Exp squared divided by expected
round(((inv_observed - inv_expected)^2 / inv_expected), 2)
sum(round(((inv_observed - inv_expected)^2 / inv_expected), 2))
```

Critical region
```{r}
# Parameters
df <- 8 # Degrees of freedom
x <- seq(0, 20, by = 0.1) # Range of x values
tail_start <- 15.51 # Starting point for the shaded tail

# Create a data frame for the chi-square distribution
data <- data.frame(
  x = x,
  y = dchisq(x, df = df)
)

# Create the plot
chisq_plot <- ggplot(data, aes(x = x, y = y)) +
  geom_line(color = ada_colour("blue"), size = 1) +
  geom_area(
    data = subset(data, x >= tail_start),
    aes(x = x, y = y),
    fill = ada_colour("blue"),
    alpha = 1
  ) +
  labs(x = "x", y = "density")

chisq_plot
```


Using the built-in test

```{r}
chisq.test(x = inv_observed, p = probabilities)
```
```{r}
# First digit probabilities
first_digit <- data.frame(
  digit = 1:9,
  probability = log10(1 + 1 / 1:9),
  position = "First"
)

# Second digit probabilities
second_digit <- data.frame(
  digit = 0:9,
  probability = sapply(0:9, function(d) sum(log10(1 + 1 / (10 * 1:9 + d)))),
  position = "Second"
)

# Third digit probabilities
third_digit <- data.frame(
  digit = 0:9,
  probability = sapply(0:9, function(d) sum(log10(1 + 1 / (10 * 10:99 + d)))),
  position = "Third"
)

# Combine data
benford_data <- bind_rows(first_digit, second_digit, third_digit)

# Create line plots
benford_digit_plot <- ggplot(benford_data, aes(x = digit, y = probability, group = position)) +
  ggplot2::theme_bw() +
  geom_line(color = ada_colour("blue")) +
  geom_point(color = ada_colour("blue")) +
  facet_wrap(~position, ncol = 1, scales = "free_x") +
  labs(
    x = "Digit",
    y = "Probability"
  ) +
  theme(legend.position = "none")

benford_digit_plot
```




```{r}
# The ordinary business-day baseline gives the expected mix of journal-entry
# types. The final two working days provide the observed counts for a focused
# period-end screen.
je_categories <- c(
  "Sales entries",
  "Purchase entries",
  "Cash receipts/payments",
  "Payroll entries",
  "Inventory and cost-of-sales entries",
  "Accruals and deferrals",
  "Reclassifications",
  "Manual adjustments"
)

je_baseline <- c(0.26, 0.22, 0.16, 0.07, 0.10, 0.08, 0.06, 0.05)
je_observed <- c(82, 74, 51, 24, 38, 63, 42, 46)
je_expected <- sum(je_observed) * je_baseline

je_summary <- data.frame(
  category = je_categories,
  observed = je_observed,
  baseline_probability = je_baseline,
  expected = je_expected,
  chi_square_component = (je_observed - je_expected)^2 / je_expected
)

je_summary

je_chi_square <- sum(je_summary$chi_square_component)
je_df <- length(je_categories) - 1
je_critical_value <- qchisq(0.95, df = je_df)
je_p_value <- pchisq(je_chi_square, df = je_df, lower.tail = FALSE)

c(
  chi_square = je_chi_square,
  degrees_of_freedom = je_df,
  critical_value = je_critical_value,
  p_value = je_p_value
)

chisq.test(x = je_observed, p = je_baseline)
```
