---
title: "Workshop Chapter 5: Regression analysis"
author: "Lucas Hoogduin"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

<!-- GENERATED FILE: edit notebooks/support/regression-analysis/support.Rmd in the private ada repository. -->


## Regression analysis

### Exercise 5.1. Before we start

To run the exercises in this workshop, we need to load a few packages.

```{r setup}
library(ggplot2) # Data visualization
theme_set(theme_gray())
library(aicpa) # Contains US SteamCo data
library(car) # Companion to Applied Regression
library(gridExtra) # Arrange multiple grid-based plots
library(tidyr) # Data tidying (reshape and transform)
library(corrplot) # Correlation matrix visualization
library(lmtest) # Diagnostic tests for linear models
library(latex2exp) # Use LaTeX expressions in plots
library(scales)
library(grid)
```

### Exercise 5.2. The US SteamCo data file

Throughout this workshop, we work with the `USSteamCo` data file, this is available in the `aicpa` package. The data file is used in the AICPA Guide to Audit Analytics, Exhibit B-3.

The `head` function shows the values of the variables (columns) in the first few observations (rows) from the dataset.

```{r}
head(USSteamCo)
```
### Exercise 5.3. Summarizing and plotting the data

The `summary` command produces summary statistics of all variables in the dataset, the length (number of data points) for character types (such as `month`) and distribution statistics for numeric types.

```{r}
summary(USSteamCo)
```

We create a dummy variable to distinguish between heating months and cooling months. Such a variable would have the value 1 in heating months and 0 in cooling months. The effect of this variable on the trajectory of the regression line is explained in Section 5.4.4.

```{r}
USSteamCo$summer <- c(0, 0, 0, 0, 1, 1, 1, 1, 1, 0, 0, 0)
```

We reformat the month variable, that has a character type, into the `date` variable.

```{r}
USSteamCo$date <- seq(as.Date("2011-01-01"),
  by = "month", length.out = 48
)
```

We split the data file into an estimation set and a hold-out set.

```{r}
USSteamCoEstim <- USSteamCo[1:36, ]
USSteamCoHold <- USSteamCo[37:48, ]
```

Figure 5.2 shows the histograms of the data in the US SteamCo estimation set.

```{r}
hist_revenue <-
  ggplot(USSteamCoEstim, aes(x = revenue)) +
  geom_histogram(binwidth = 4000000, fill = "#00338D") +
  scale_x_continuous(labels = label_comma())

hist_production <-
  ggplot(USSteamCoEstim, aes(x = production)) +
  geom_histogram(binwidth = 50000, fill = "#00338D") +
  scale_x_continuous(labels = label_comma()) +
  theme(plot.margin = unit(c(0, 1, 0, 0), "cm")) # top, right, bottom, left

hist_coolDD <-
  ggplot(USSteamCoEstim, aes(x = coolDD)) +
  geom_histogram(binwidth = 50, fill = "#00338D") +
  scale_x_continuous(labels = label_comma())

hist_heatDD <-
  ggplot(USSteamCoEstim, aes(x = heatDD)) +
  geom_histogram(binwidth = 100, fill = "#00338D") +
  scale_x_continuous(labels = label_comma())

histograms_plot <- grid.arrange(hist_revenue, hist_production, hist_coolDD, hist_heatDD, ncol = 2)
```

Figure 5.3 shows the time-series plots for the same dataset.

```{r}
# Calculate the range of revenue and production
rev_range <- range(USSteamCoEstim$revenue, na.rm = TRUE)
prod_range <- range(USSteamCoEstim$production, na.rm = TRUE)

# Create the plot
time_series_plot <- ggplot(data = USSteamCoEstim, aes(x = date)) +
  # First time series (revenue)
  geom_line(aes(y = revenue), colour = "#00338D") +
  # Second time series (production), rescaled to match revenue range
  geom_line(aes(y = scales::rescale(production, to = rev_range)),
    colour = "#BC204B"
  ) +
  scale_y_continuous(
    name = "Revenue ($)",
    labels = scales::label_comma(),
    sec.axis = sec_axis(
      ~ scales::rescale(., from = rev_range, to = prod_range),
      name = "Production (x 1000 lb)",
      labels = scales::label_comma()
    )
  ) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y.left = element_text(color = "#00338D"),
    axis.title.y.right = element_text(color = "#BC204B")
  )
time_series_plot
```


Figure 5.1 shows a scatter plot of the independent variable production ($x$-axis) and the dependent variable revenue ($y$-axis) for the Case: US SteamCo. The plot is an easy means to get a feel for the relationship.

```{r}
scatter_plot <- ggplot(USSteamCoEstim, aes(x = production, y = revenue)) +
  geom_point(color = "#00338D") +
  scale_x_continuous(name = "Production (x 1000 lb)", labels = scales::label_comma()) +
  scale_y_continuous(name = "Revenue ($)", labels = scales::label_comma())
scatter_plot
```

The enhanced scatter plot in Figure 5.4 is obtained by running the following code.

```{r}
scatterplot(revenue ~ production,
  data = USSteamCoEstim,
  smooth = FALSE
)
```

### Exercise 5.4. The base model `mod.0`

Figure 5.4 shows that there is a positive (increasing) relationship: as production increases, so does revenue.

We use the function `lm` (for linear model) to estimate the coefficient values of $\beta_0$ and $\beta_1$. This is performed on the data of the estimation set `USSteamCoEstim`, as explained in Section 5.1.1. A brief summary of the model is obtained with the function `brief`.

```{r}
ussteamco.mod.0 <- lm(revenue ~ production, data = USSteamCoEstim)
brief(ussteamco.mod.0)
```
The estimate of the intercept $\beta_0$, $b_0 = 4,897,663$, and the estimate of the slope $\beta_1$, $b_1 = 18.99$.

### Exercise 5.5. The full summary of `mod.0`

A full summary of `mod.0` can be obtained with the `summary` function.

```{r}
summary(ussteamco.mod.0)
```
In addition to the statistics produced by `brief`, it provides insight into the residuals (minimum, first quartile, median, third quartile, and maximum), the $t$ and $p$ values of each of the coefficients and their significance, the adjusted $r^2$, and the $F$ statistic with its $p$ value.

### Exercise 5.6. Correlations and correlogram

The correlations in Table 5.3 are calculated using the `cor` function.

```{r}
(cor_ussteam <- cor(USSteamCoEstim[, 2:5]))
```
The correlogram in Figure 5.5 is produced by

```{r echo=FALSE}
corrplot(cor_ussteam,
  tl.col = "black",
  col = colorRampPalette(c("#bc204b", "white", "#00348d"))(100),
  col.lim = c(-1, 1)
)
```

### Exercise 5.7. Modelling strategies

The `step` function can be used to implement both backward and forward methods. The `direction` argument specifies whether the stepwise algorithm should proceed forward (`forward`), backward (`backward`), or both (`both`). The `scope` argument specifies the range of models to be searched, whereas the `k` argument may be used to use BIC instead of the default AIC as a criterion.

We start with the forward method. From a model that only uses the constant (`revenue $\sim$ 1`), we add variables one by one. The variable that the largest effect on the criterion, in this example, the default AIC, is selected for inclusion in the model.

```{r}
model_forward <- lm(revenue ~ 1, data = USSteamCoEstim)
model_forward <- step(model_forward,
  direction = "forward",
  scope = formula(~ production + coolDD + heatDD)
)
```
Starting with AIC = 1114.66 of a model with only a constant, `R` evaluates the inclusion of each of the available independent variables and their effect on the resulting AIC. In the first step, it appears that including `heatDD` has the best effect on AIC, as it reduces from 1114.66 to 1096.56.

Using this value of AIC = 1096.56 as the new starting value, `R` evaluates if inclusion of any of the remaining variables has a favorable effect. It finds that including `production` can further reduce the AIC to 1046.11.

In the last step, we see that including `coolDD` does not further reduce the AIC, leading to the final model `revenue $\sim$ heatDD + production`. The summary of this model is:

```{r}
summary(model_forward)
```

Next, we show an example of the backward method, starting with the full model and eliminating variables one by one.

```{r}
model_backward <- lm(revenue ~ production + coolDD + heatDD,
  data = USSteamCoEstim
)
model_backward <- step(model_backward,
  direction = "backward"
)
summary(model_backward)
```

We observe that this strategy results into the exact same model.

To run the stepwise procedure in both directions, we use the following code.

```{r}
fit_both <- lm(revenue ~ 1, data = USSteamCoEstim)
fit_both <- step(fit_both,
  direction = "both",
  scope = formula(~ production + coolDD + heatDD)
)
summary(fit_both)
```

### Exercise 5.8. Multiple independent variables `mod.1`

We now regress $y$ = `revenue` on $x_1$ = `production`, $x_2$ = `coolDD`, and $x_3$ = `heatDD`.

```{r}
ussteamco.mod.1 <- lm(revenue ~ production + coolDD + heatDD,
  data = USSteamCoEstim
)
summary(ussteamco.mod.1)
```

Adding the additional variables `coolDD` and `heatDD` has improved the fit of the model. The residual standard deviation decreased from 4,112,487 for `mod.0` to 1,980,698 for `mod.1`. Similarly, $r^2$ improved from 0.397 to 0.868. These metrics suggest that the inclusion of relevant variables enhances the explanatory power of the model, demonstrating the importance of carefully chosen criteria in stepwise selection methods.

## Exercise 5.9. Interactions

In Section 5.4.4 we developed `mod.2`, that incorporates two separate regression lines. In addition to the three independent variables, we also account for their interactions with the dummy variable `summer`.

The scatterplot in Figure 5.6 is prepared using

```{r}
scat_rev2 <- scatterplot(revenue ~ production | summer,
  data = USSteamCoEstim,
  id = list(n = 8, cex = 0.8),
  regLine = TRUE,
  smooth = FALSE
)
```

```{r}
# Picture in ADA

# Ensure summer is a factor
USSteamCoEstim$summer_fact <- factor(USSteamCoEstim$summer, levels = c(0, 1), labels = c("No", "Yes"))

# Define custom colors
summer_colors <- c("No" = "#00338D", "Yes" = "#BC204B")

# Base plot with manual color scale and separate regression lines
p <- ggplot(USSteamCoEstim, aes(x = production, y = revenue, color = summer_fact)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(values = summer_colors) +
  scale_x_continuous(labels = label_comma()) +
  scale_y_continuous(labels = label_comma()) +
  labs(x = "Production", y = "Revenue", color = "Summer")

p
```


Note the special operator `|`, that can be expressed as 'given the value of'. To call this relationship, we use

```{r}
ussteamco.mod.2 <- lm(
  revenue ~ production * summer +
    coolDD * summer +
    heatDD * summer,
  data = USSteamCoEstim
)
summary(ussteamco.mod.2)
```
The formula specified in the `lm` function, `revenue $\sim$ production * summer + coolDD * summer + heatDD * summer` can also be written as `revenue $\sim$ (production + coolDD + heatDD) * summer`.

As a result, this further improved the fit of the model. The residual standard deviation decreased to 1,680,711, and $r^2$ improved to 0.917.

## Exercise 5.10. Time lag

The cross-correlation plot in Figure 5.7 is created by running:

```{r}
with(
  USSteamCoEstim,
  ccf(
    x = production,
    y = revenue,
    lag.max = 4
  )
)
```


## Exercise 5.11. Residual plots

The residual plots in Figure 5.9 are obtained by:

```{r}
# Add Pearson residuals to the data
USSteamCoEstim$resid <- residuals(ussteamco.mod.2, type = "pearson")

# Extract predictor names
predictors <- all.vars(formula(ussteamco.mod.2))[-1]

# Function to create residual plot
make_resid_plot <- function(data, var) {
  n_unique <- length(unique(data[[var]]))

  p <- ggplot(data, aes(x = .data[[var]], y = resid)) +
    geom_point(color = "#00338D", alpha = 0.7) +
    geom_hline(yintercept = 0, linetype = "dotted") +
    scale_x_continuous(labels = label_comma()) +
    scale_y_continuous(labels = label_comma()) +
    labs(x = var, y = "")

  if (n_unique > 2) {
    p <- p +
      geom_smooth(
        method = "lm",
        formula = y ~ poly(x, 2),
        se = FALSE,
        color = "#E36877"
      )
  } else {
    p <- p +
      geom_smooth(
        method = "lm",
        formula = y ~ x,
        se = FALSE,
        color = "#E36877"
      )
  }

  p
}

# Create residual plots for each predictor
plots <- lapply(predictors, function(var) {
  make_resid_plot(USSteamCoEstim, var)
})

# Fitted values vs residuals plot
fitted_data <- data.frame(
  fitted = fitted(ussteamco.mod.2),
  resid = USSteamCoEstim$resid
)

fitted_plot <- ggplot(fitted_data, aes(x = fitted, y = resid)) +
  geom_point(color = "#00338D", alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_smooth(
    method = "lm",
    formula = y ~ poly(x, 2),
    se = FALSE,
    color = "#E36877"
  ) +
  scale_x_continuous(labels = label_comma()) +
  scale_y_continuous(labels = label_comma()) +
  labs(x = "Fitted Values", y = "")

# Top grid: 2x2 layout of the first 4 residual plots
top <- arrangeGrob(
  grobs = plots[1:4],
  ncol = 2
)

# Combine top and bottom
combined <- arrangeGrob(
  top,
  fitted_plot,
  ncol = 1,
  heights = c(2, 1)
)

# Display
grid.newpage()
grid.draw(combined)
```



## Exercise 5.12. Influence statistics

Figure 5.10 provides the Influence index plots. From bottom to top: Leverage points, regression outliers, their respective Bonferroni $p$ values and influential observations.

```{r}
influenceIndexPlot(ussteamco.mod.2)
```

## Exercise 5.13. Leverage points

We obtain the hat-values of a model by running the `hatvalues` function on it.

```{r}
hatvals.mod.2 <- hatvalues(ussteamco.mod.2)
```

The three largest hat-values are then summarized by

```{r}
hatvals.mod.2[order(hatvals.mod.2, decreasing = TRUE)[1:3]]
```

The sum of the hat-values is equal to $k + 1$.

```{r}
sum(hatvals.mod.2)
```

## Exercise 5.14. Regression outliers

Figure 5.11 provides the QQ plot for \ttblue{ussteamco.mod.2}, with the three largest Studentized residuals identified.

```{r}
qqPlot(ussteamco.mod.2, id = list(n = 3))
```

## Exercise 5.15. Influential observations

We obtain the Cook's distances of a model by running the `cooks.distance` function on it.

```{r}
cooks.mod.2 <- cooks.distance(ussteamco.mod.2)
```

The three largest Cook's distances are then summarized by

```{r}
cooks.mod.2[order(cooks.mod.2, decreasing = TRUE)[1:3]]
```

## Exercise 5.16. Infuence plot

Figure 5.12 shows the hat-values, Studentized residuals, and Cook’s distances for `ussteamco.mod.2`. The size of the circles is proportional to the Cook’s distance.

```{r}
influencePlot(ussteamco.mod.2, id = list(n = 3, cex = 0.8))
```

## Exercise 5.17. Winsorizing observation # 22

Define the new estimation set. We replace the original value of observation 22 with its winsorized value.

```{r}
USSteamCoEstim2 <- USSteamCoEstim
p5 <- quantile(ussteamco.mod.2$residuals, 0.05)
fit_22 <- fitted(ussteamco.mod.2)[22]
USSteamCoEstim2$revenue[22] <- fit_22 + p5
```

We then fit a new linear model.

```{r}
ussteamco.mod.3 <- lm(revenue ~ (production + heatDD + coolDD) * summer,
  data = USSteamCoEstim2
)

brief(ussteamco.mod.3)
```

## Exercise 5.18. Multicollinearity

To assess multicollinearity for a model without interaction terms, such as `mod.1`, we use the standard `vif()` function.

```{r}
vif(ussteamco.mod.1)
```

A common rule of thumb is that VIF values greater than 10 indicate substantial multicollinearity. Since none of the VIF values exceeds this threshold, we conclude that there is no evidence of serious variance inflation in `mod.1`.

For models that include interaction terms, such as `mod.3`, we use the `vif()` function with the additional option `type = "predictor"`. This reports the Generalized Variance Inflation Factor (GVIF) and its adjusted value, \(GVIF^{1/(2df)}\).

```{r}
vif(ussteamco.mod.3, type = "predictor")
```

The adjusted GVIF places predictors with different degrees of freedom on a comparable scale. For a predictor with one degree of freedom, the adjusted GVIF equals the square root of the ordinary VIF. Consequently, the conventional VIF threshold of 10 corresponds to an adjusted GVIF of approximately \(\sqrt{10} \approx 3.16\). Since none of the reported values exceeds 3.16, we conclude that `mod.3` does not exhibit problematic multicollinearity.

## Exercise 5.19. Variance and Total Sum of Squares

Table 5.5 provides the analysis of variance for `mod.0`. We apply the `anova()` command to perform this analysis.

```{r}
anova(ussteamco.mod.0)
```

The variance of $y$, $s_y^2 = \frac{\sum(y_i - \bar{y})^2}{n - 1}$, is obtained from

```{r}
var(USSteamCoEstim$revenue)
```

Total Sum of Squares $\sum(y_i - \bar{y})^2$ is calculated as

```{r}
mean_y <- mean(USSteamCoEstim$revenue)
variation <- USSteamCoEstim$revenue - mean_y
squared_variation <- variation^2
(total_sum_of_squares <- sum(squared_variation))
```

Total Sum of Squares is the sum of the elements in the `Sum Sq` column of the Analysis of Variance Table.

```{r}
anova.mod.0 <- anova(ussteamco.mod.0)
anova.mod.0$`Sum Sq`[1] + anova.mod.0$`Sum Sq`[2]
```

When we divide Total Sum of Squares by $n - 1$, we get the variance of $y$.

```{r}
total_sum_of_squares / 35
```

This shows how variance is linked to variation.

## Exercise 5.20. Anova Type I and Type II

We now turn to a model with multiple independent variables, for example, `mod.1`, introduced in Section 5.4.3.

```{r}
(anova.mod.1 <- anova(ussteamco.mod.1))
```

Declaring the independent variables in a different order shows that the calculation of the sum of squares is in sequential order.

```{r}
ussteamco.mod.1b <- lm(revenue ~ heatDD + coolDD + production,
  data = USSteamCoEstim
)
(anova.mod.1b <- anova(ussteamco.mod.1b))
```
To address this issue, an alternative version of the analysis of variance, referred to as Type II tests, is used. Instead of measuring the contribution of each independent variable sequentially, the analysis shows the contribution of each independent variable if it was the last independent variable added to the model. The analysis is obtained using the command `Anova` (with capital A) from the `car` package.

```{r}
Anova(ussteamco.mod.1)
```

We see that the marginal effect of adding `production`, assuming `coolDD` and `heatDD` have already been entered into the model, is 1.8260e+14. This is equal to the result obtained from the earlier `anova` command on `mod.1b`. We also see that the added value of entering `coolDD` into the model is limited: its sum of squares is much lower than that of the other two variables, and its related $F$ value is not significant.

We have no clear preference for either `anova` or `Anova`, they both have their merits. `anova` decomposes TSS, but that decomposition is dependent on the order in which variables are entered into the equation. `Anova` shows the marginal effect of each independent variable, but these marginal effects do not add to TSS.

## Exercise 5.21. Anova with interactions

The analysis of variance of `mod.3` is as follows:

```{r}
(anova.mod.3 <- anova(ussteamco.mod.3))
```

## Exercise 5.22. Comparing models

We compare the performance of models with four different statistics, the Multiple $r^2$, Adjusted $r^2$ (both obtained from the `summary` function), and the AIC and BIC values. These last two can be called with their own respective commands.

```{r}
AIC(ussteamco.mod.0)
AIC(model_backward)
AIC(ussteamco.mod.1)
AIC(ussteamco.mod.2)
```
The AIC values shown above can be compared directly because all four models are fitted to the same response data (`USSteamCoEstim`). Model `mod.3` is fitted after winsorizing observation 22 and therefore uses a modified response variable. Since the likelihood is computed from the observed responses, the AIC (and BIC) of `mod.3` is not directly comparable with those of the earlier models. Instead, `mod.3` should be regarded as a sensitivity analysis demonstrating the effect of mitigating the influence of an outlying observation.

The AIC value of the `model_backward` does not quite match the output of the `step` function. This can be explained as follows.

The Akaike Information Criterion (AIC) provides a penalized measure of model fit, defined as
$\text{AIC} = -2 \log \hat{L} + 2k$,
where $\log \hat{L}$ is the maximized log-likelihood of the model and $k$ is the number of estimated parameters.

For a linear regression model with normally distributed errors, this log-likelihood takes the form
$\log \hat{L} = -\frac{n}{2} \left[ \log(2\pi) + 1 + \log\left(\frac{\text{SSE}}{n}\right) \right]$,
leading to the AIC expression
$\text{AIC} = n \left[\log(2\pi) + 1 + \log\left(\frac{\text{SSE}}{n}\right)\right] + 2k$.
In contrast, the `step()` function in R reports AIC values based on a deviance-like approximation, that omits constant terms unrelated to model comparison:
$\text{AIC}_{\text{step}} = n \log\left(\frac{\text{SSE}}{n}\right) + 2k$.
The difference between the two expressions is a constant shift of
$n \cdot (\log(2\pi) + 1) + 2$.
The term $n \cdot (\log(2\pi) + 1)$ reflects the part of the log-likelihood that depends only on the sample size and the assumption of normally distributed errors.
The additional +2 arises because R’s AIC() function evaluates the likelihood based on a model with an estimated variance parameter $\sigma^2$, introducing one more estimated parameter than is typically counted in the model formula (e.g., intercept and slopes). This shifts the effective count of parameters k by 1 and contributes an extra penalty of 2 in the AIC expression.

Similarly, the BIC values are obtained as follows.
```{r}
BIC(ussteamco.mod.0)
BIC(model_backward)
BIC(ussteamco.mod.1)
BIC(ussteamco.mod.2)
```

The same consideration applies to the BIC. The BIC values of `mod.0`, `model_backward`, `mod.1`, and `mod.2` may be compared directly because they are based on the same response data. The BIC of `mod.3` should not be interpreted as evidence that the winsorized model is better or worse than the previous models, because it is based on a different dataset.

## Exercise 5.23. Testing normality

We can assess normality issues with a histogram or with a QQ plot. The histogram in Figure 5.14 is obtained by:

```{r}
# Calculate normalized residuals
nres <- ussteamco.mod.3$residuals / sd(ussteamco.mod.3$residuals)
model3 <- as.data.frame(nres)
hist_stand_res <-
  ggplot(model3, aes(nres)) +
  geom_histogram(aes(y = after_stat(density)),
    breaks = seq(-2.55, 1.65, by = 0.7),
    fill = "#00338D"
  ) +
  stat_function(fun = dnorm, args = list(
    mean = mean(model3$nres),
    sd = sd(model3$nres)
  ))
hist_stand_res
```

The QQ plot is produced by:

```{r}
qqPlot(ussteamco.mod.3)
```


The Shapiro-Wilk normality test is executed through the `shapiro.test` function from the `stats` package.

```{r}
shapiro.test(ussteamco.mod.3$residuals)
```

## Exercise 5.24. Testing heteroskedasticity

To test the asumption of constant variance, we can employ visual means, such as the `residualPlots` in Figure 5.15. We can construct it as follows.

```{r}
residualPlots(ussteamco.mod.3,
  layout = c(2, 3),
  quadratic = FALSE, linear = TRUE
)
```
The formal statistical test is the Breusch-Pagan test.

```{r}
bptest(ussteamco.mod.3)
```

## Exercise 5.25. Testing autocorrelation

We use the `pacf` plot from Figure 5.17 to see if autocorrelation is present, and if so, which order is significant.

```{r}
pacf(ussteamco.mod.3$residuals)
```

The Breusch-Godfrey test is executed through `bgtest`.

```{r}
bgtest(ussteamco.mod.3, order = 3, type = "Chisq")
```

The $p$ value of the test statistic is 0.06737. Therefore, we reject the null hypothesis at levels below 0.06737, suggesting marginal evidence of autocorrelation in the residuals.

## Exercise 5.26. First-order autocorrelation correction

A practical way to address first-order autocorrelation is to apply an AR(1) transformation. This method estimates the autocorrelation parameter and transforms the regression model by subtracting a multiple of the lagged value from the dependent variable and from each predictor. The transformed model can then be estimated by ordinary least squares.

We first estimate an AR(1) model on the residuals from `mod.3`.

```{r}
(rho <- arima(ussteamco.mod.3$residuals,
  order = c(1, 0, 0)
)$coef[1])
```

Then we transform all variables using the estimated \(\rho\). When the model includes interaction terms, we first construct the interaction terms in their original form and then transform those interaction terms. We should not multiply separately transformed variables.

```{r}
# Create lagged variables, dropping the first observation
n <- nrow(USSteamCoEstim2)

# Current period t and lagged period t - 1
trans <- USSteamCoEstim2[-1, ]
lag <- USSteamCoEstim2[-n, ]

# Construct original interaction terms first
trans$prod_summer <- trans$production * trans$summer
trans$heat_summer <- trans$heatDD * trans$summer
trans$cool_summer <- trans$coolDD * trans$summer

lag$prod_summer <- lag$production * lag$summer
lag$heat_summer <- lag$heatDD * lag$summer
lag$cool_summer <- lag$coolDD * lag$summer

# Transform dependent variable
trans$revenue_adj <- trans$revenue - rho * lag$revenue

# Transform independent variables
trans$production_adj <- trans$production - rho * lag$production
trans$heatDD_adj <- trans$heatDD - rho * lag$heatDD
trans$coolDD_adj <- trans$coolDD - rho * lag$coolDD
trans$summer_adj <- trans$summer - rho * lag$summer

# Transform interaction terms
trans$prod_summer_adj <- trans$prod_summer - rho * lag$prod_summer
trans$heat_summer_adj <- trans$heat_summer - rho * lag$heat_summer
trans$cool_summer_adj <- trans$cool_summer - rho * lag$cool_summer
```

We fit a new model using the transformed variables.

```{r}
ussteamco.mod.4 <- lm(
  revenue_adj ~
    production_adj + heatDD_adj + coolDD_adj +
    summer_adj + prod_summer_adj +
    heat_summer_adj + cool_summer_adj,
  data = trans
)

summary(ussteamco.mod.4)
```

This transformation follows $y_t - \rho y_{t-1} = \beta_0(1-\rho) + \sum_j \beta_j(x_{j,t} - \rho x_{j,t-1}) + u_t$.

For interaction terms, the interaction is treated as a predictor in its own right. Thus, if \(z_t=x_t s_t\), the transformed interaction is $z_t-\rho z_{t-1} = x_t s_t-\rho x_{t-1}s_{t-1}$.

It is not obtained by multiplying the transformed variables. The first observation is discarded because no lagged value is available. After fitting the transformed model, we check whether autocorrelation is reduced by observing the spikes in the `pacf` plot and running the Breusch-Godfrey test.

```{r}
pacf(ussteamco.mod.4$residuals)

bgtest(ussteamco.mod.4, order = 3)
```
## Exercise 5.27. Constructing a refined model

When a model includes interaction terms, it is important to respect the hierarchical structure—interaction terms should not be included without their corresponding main effects. The base `R` function `step()` enforces this principle automatically by removing interaction terms before dropping associated main effects. Preserving model hierarchy is essential for ensuring interpretability and theoretical coherence, especially in models intended for inference.

```{r}
ussteamco.mod.5 <- lm(revenue ~ (production + coolDD + heatDD) * summer,
  data = USSteamCoEstim2
)
ussteamco.mod.5 <- step(ussteamco.mod.5,
  direction = "backward"
)
```

## Exercise 5.28. Coefficient test of the refined model

The summary of this model is:

```{r}
options(scipen = 999)
options(digits = 9)
summary(ussteamco.mod.5)
```

## Exercise 5.29. Re-running the model assumption tests

We re-run the model assumption tests.

```{r}
bptest(ussteamco.mod.5)
bgtest(ussteamco.mod.5, order = 3, type = "Chisq")
shapiro.test(ussteamco.mod.5$residuals)
```

## Exercise 5.30. Testing significance

The $F$ statistic used for testing the significance of the model as a whole can also be obtained from the \ttblue{summary()} table.

```{r}
summary(ussteamco.mod.5)
```

### Exercise 5.31. Visualizing the uncertainty in the model parameters

First, we set the values of the parameters. Play around with these to see the effect on the ultimate graph!

```{r}
set.seed(123) # Set seed for reproducibility
beta_0 <- 2 # True intercept
beta_1 <- 1.5 # True slope
n <- 11 # Number of points to sample
x_min <- 0 # Minimum value of x
x_max <- 10 # Maximum value of x
x_values <- seq(x_min, x_max, length.out = n) # x values modelled
sigma <- 5 # Standard deviation of the noise
rep <- 200 # Number of replications
alpha <- 0.05 # Significance level for 95% confidence interval
x_extended_min <- x_min - (x_max - x_min) / 2 # Extended x range minimum
x_extended_max <- x_max + (x_max - x_min) / 2 # Extended x range maximum
x_extended <- seq(x_extended_min, x_extended_max, length.out = 100) # Plot range
```

We first calculate the true regression line, along with its confidence interval.

```{r}
# Create a data frame for the true line over the extended range
true_line <- data.frame(x = x_extended, y = beta_0 + beta_1 * x_extended)

# Calculate the standard error of the true regression line
mean_x <- mean(x_values) # Mean of the x values
S_xx <- sum((x_values - mean_x)^2) # Sum of squares of x deviations
se_fit_true <- sigma * sqrt(1 / n + (x_extended - mean_x)^2 / S_xx) # Standard error formula for regression

# Calculate the theoretical confidence intervals based on the true model
t_value <- qt(1 - alpha / 2, df = n - 2) # Critical t-value for 95% CI
ci_upper_true <- (beta_0 + beta_1 * x_extended) + t_value * se_fit_true # Upper bound
ci_lower_true <- (beta_0 + beta_1 * x_extended) - t_value * se_fit_true # Lower bound

# Create a data frame for the confidence interval bounds based on the true regression line
ci_bounds_true <- data.frame(x = x_extended, ci_lower = ci_lower_true, ci_upper = ci_upper_true)
```

Next, we sample $n$ points from a normal distribution, to generate $y$ values for each $x$ value and fit the corresponding regression line. We repeat this process `rep` times.

```{r}
# Initialize a data frame to store the simulated regression lines
simulated_lines <- data.frame()

# Repeat the process `rep` times to simulate multiple regression lines
for (i in 1:rep) {
  # Sample n points from the relationship y = beta_0 + beta_1 * x + epsilon
  epsilon <- rnorm(n, mean = 0, sd = sigma) # Random noise
  y_values <- beta_0 + beta_1 * x_values + epsilon

  # Fit a linear model
  model <- lm(y_values ~ x_values)

  # Predict fitted values for the extended range of x
  predicted_values <- predict(model, newdata = data.frame(x_values = x_extended))

  # Store the fitted line over the extended x range
  simulated_lines <- rbind(
    simulated_lines,
    data.frame(
      x = x_extended,
      y = predicted_values,
      rep = i
    )
  )
}
```

All the elements created earlier are now plotted on a single canvas.

```{r}
regression_lines_plot <- ggplot() +
  geom_line(
    data = simulated_lines, aes(x = x, y = y, group = rep),
    color = "#00338D", alpha = 0.1
  ) + # Plot the simulated regression lines
  geom_line(data = true_line, aes(x = x, y = y), color = "#E36877", linewidth = .5) + # Plot the true line
  geom_line(
    data = ci_bounds_true, aes(x = x, y = ci_lower),
    color = "#E36877", linewidth = .5, linetype = "solid"
  ) + # Lower bound (true line)
  geom_line(
    data = ci_bounds_true, aes(x = x, y = ci_upper),
    color = "#E36877", linewidth = .5, linetype = "solid"
  ) + # Upper bound (true line)
  labs(
    title = paste(rep, " Regression Lines with Confidence Interval Based on True Model"),
    x = "x", y = "y"
  ) +
  theme_minimal() +
  xlim(x_extended_min, x_extended_max) # Extend the x-axis limits
regression_lines_plot
```

The resulting graph is shown in Figure 5.20.

## Exercise 5.32. Confidence and prediction intervals

Figure 5.18 shows a regression line with a dotted confidence interval and a shaded prediction interval. It is created with the following code.

```{r}
# Calculate the standard error for prediction interval
alpha_pred <- 0.01 # For 99% prediction interval
t_value_pred <- qt(1 - alpha_pred / 2, df = n - 2)
se_pred <- sigma * sqrt(1 + 1 / n + (x_extended - mean_x)^2 / S_xx)

# Calculate 99% prediction interval bounds
pi_upper_true <- (beta_0 + beta_1 * x_extended) + t_value_pred * se_pred
pi_lower_true <- (beta_0 + beta_1 * x_extended) - t_value_pred * se_pred

pi_bounds_true <- data.frame(
  x = x_extended,
  pi_lower = pi_lower_true,
  pi_upper = pi_upper_true
)

intervals_plot <- ggplot() +
  geom_ribbon(
    data = pi_bounds_true,
    aes(x = x, ymin = pi_lower, ymax = pi_upper),
    fill = "#00338D", alpha = 0.15
  ) + # 99% prediction interval as a ribbon
  geom_line(
    data = ci_bounds_true,
    aes(x = x, y = ci_lower),
    color = "#00338D", linewidth = .5, linetype = "dashed"
  ) + # 95% CI lower
  geom_line(
    data = ci_bounds_true,
    aes(x = x, y = ci_upper),
    color = "#00338D", linewidth = .5, linetype = "dashed"
  ) + # 95% CI upper
  geom_line(
    data = true_line,
    aes(x = x, y = y),
    color = "#00338D", linewidth = 1
  ) + # True regression line
  labs(
    x = "x", y = "y"
  ) +
  xlim(x_extended_min, x_extended_max)
intervals_plot
```


## Exercise 5.33. Creating expectations for the hold-out period

The function `predict` creates expectations, when data from `newdata` are evaluated with a certain `object`. The optional `interval` is used to add bounds for a prediction interval, or, alternatively, a confidence interval. The `level` specifies the confidence level of the interval.

```{r echo=FALSE}
(predictions <- predict(
  object = ussteamco.mod.5,
  newdata = USSteamCoHold,
  interval = "prediction",
  level = 0.99
))
```
Compare the recorded revenue with the expectations.

```{r}
comparison <- cbind(
  Month = rownames(USSteamCoHold),
  Recorded = USSteamCoHold$revenue,
  Lower = round(predictions[, "lwr"], 0),
  Expected = round(predictions[, "fit"], 0),
  Upper = round(predictions[, "upr"], 0)
)

comparison <- as.data.frame(comparison)

comparison$Recorded <- as.numeric(comparison$Recorded)
comparison$Expected <- as.numeric(comparison$Expected)
comparison$Lower <- as.numeric(comparison$Lower)
comparison$Upper <- as.numeric(comparison$Upper)

# Difference = Recorded - Expected
comparison$Difference <- comparison$Recorded - comparison$Expected

comparison
sum(comparison$Recorded)
sum(comparison$Expected)
sum(comparison$Difference)
```


## Exercise 5.34. Plotting expectations against recorded values

In Figure 5.19 we have plotted expectations against the recorded values of revenue, indicating which individual observations fall outside of the 99% prediction interval. The following code is used to create it.

```{r}
# Convert predictions to a data frame and add the date column
pred_df <- data.frame(
  date = USSteamCoHold$date,
  recorded = USSteamCoHold$revenue,
  lwr = predictions[, "lwr"],
  fit = predictions[, "fit"],
  upr = predictions[, "upr"]
)
pred_df$outside <- with(pred_df, recorded < lwr | recorded > upr)

expectations_plot <- ggplot(pred_df, aes(x = date)) +
  # Plot the predicted values as a line
  geom_line(aes(y = fit, color = "Expectation"), linewidth = 1) +
  # Add the prediction interval as a shaded area
  geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "#00338D", alpha = 0.2) +
  # Plot the actual values, with conditional coloring for outside interval points
  geom_line(aes(y = recorded, color = "Recorded"), linewidth = 1) +
  # Highlight points outside interval in red
  geom_point(data = subset(pred_df, outside), aes(y = recorded), color = "#E36877", size = 2) +
  # Define color for predicted and actual values
  scale_color_manual(
    values = c("Expectation" = "#00338D", "Recorded" = "#E36877"),
    name = ""
  ) +
  labs(
    x = "Month in 2014",
    y = "Revenue ($)"
  ) +
  scale_x_date(
    date_breaks = "1 month", # Adjusts to one gridline per month
    date_labels = "%b" # Optional: shows abbreviated month names
  ) +
  scale_y_continuous(labels = label_comma()) +
  theme(legend.position = "bottom")

expectations_plot
```

## Exercise 5.35. Combining 12 monthly predictions

We use Equations 5.29 and 5.30 to construct a prediction interval for total revenue for the year under audit. The annual expectation is the sum of the 12 monthly expectations.

The prediction standard error of the annual total is calculated directly from the design matrix for the hold-out observations. This accounts for the covariance between the monthly expectations caused by using the same estimated regression coefficients.

```{r}
# Residual degrees of freedom
df_res <- ussteamco.mod.5$df.residual

# Monthly predictions
monthly_predictions <- pred_df$fit

# Annual expectation
annual_prediction <- sum(monthly_predictions)

# Construct the design matrix for the hold-out observations
X_hold <- model.matrix(
  formula(ussteamco.mod.5),
  data = USSteamCoHold
)

# Vector of ones for summing the 12 monthly predictions
one <- rep(1, nrow(X_hold))

# Variance component from uncertainty in the estimated coefficients
var_mean_annual <- as.numeric(
  t(one) %*% X_hold %*% vcov(ussteamco.mod.5) %*% t(X_hold) %*% one
)

# Variance component from future residual variation
sigma2 <- summary(ussteamco.mod.5)$sigma^2
var_future_annual <- nrow(X_hold) * sigma2

# Prediction variance and prediction standard error of the annual total
annual_var <- var_mean_annual + var_future_annual
annual_se <- sqrt(annual_var)

# t-score for a 99% prediction interval
t_score <- qt(0.995, df = df_res)

# 99% prediction interval for the annual total
annual_lower <- annual_prediction - t_score * annual_se
annual_upper <- annual_prediction + t_score * annual_se

# Print results
cat("Variance from coefficient uncertainty:", var_mean_annual, "\n")
cat("Variance from future residual variation:", var_future_annual, "\n")

cat("99% Prediction Interval for Annual Revenue:\n")
cat("Lower Bound:", round(annual_lower, 0), "\n")
cat("Annual Prediction:", round(annual_prediction, 0), "\n")
cat("Upper Bound:", round(annual_upper, 0), "\n")
cat("Annual prediction standard error:", round(annual_se, 0), "\n")
```

The first variance component, `var_mean_annual`, captures the uncertainty in the estimated regression coefficients. The second component, `var_future_annual`, captures the future residual variation of the 12 monthly observations. The latter is calculated under the assumption that the future residual errors are uncorrelated.

Adjust the March prediction based on the corroborated evidence for the winter conditions and disruptions.

```{r}
storm_adjustment <- 8000000

pred_df$fit_adj <- pred_df$fit
pred_df$lwr_adj <- pred_df$lwr
pred_df$upr_adj <- pred_df$upr

march_row <- format(pred_df$date, "%b") == "Mar"

pred_df$fit_adj[march_row] <- pred_df$fit[march_row] - storm_adjustment
pred_df$lwr_adj[march_row] <- pred_df$lwr[march_row] - storm_adjustment
pred_df$upr_adj[march_row] <- pred_df$upr[march_row] - storm_adjustment

pred_df$outside_adj <- with(pred_df, recorded < lwr_adj | recorded > upr_adj)
pred_df
```
