---
title: "Workshop Chapter 5: Regression analysis"
author: "Lucas Hoogduin"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

<!-- GENERATED FILE: edit notebooks/support/regression-analysis/support.Rmd in the private ada repository. -->


## Regression analysis

### Exercise 5.1. Before we start

This first exercise prepares the regression-analysis workspace. The practical goal is to load the tools used for data inspection, model fitting, plotting, and diagnostics before any audit model is built.

To run the exercises in this workshop, we need to load a few packages.

```{r setup}
library(ggplot2) # Data visualization
ada_colour <- function(name) {
  palette <- c(
    blue = "#3c4b99",
    mediumblue = "#6098d2",
    lightblue = "#62b8e7",
    red = "#c93f55",
    yellow = "#eacc62",
    green = "#469d76",
    purple = "#924099",
    lightpurple = "#df9ed4",
    blue50 = "#d3d8ef",
    lightpink = "#e36877"
  )
  unname(palette[[name]])
}
theme_set(ggplot2::theme_bw())
library(aicpa) # Contains US SteamCo data
library(car) # Companion to Applied Regression
library(gridExtra) # Arrange multiple grid-based plots
library(tidyr) # Data tidying (reshape and transform)
library(corrplot) # Correlation matrix visualization
library(lmtest) # Diagnostic tests for linear models
library(latex2exp) # Use LaTeX expressions in plots
library(scales)
library(grid)
```

### Exercise 5.2. The US SteamCo data file

Throughout this workshop, we work with the `USSteamCo` data file, this is available in the `aicpa` package. The data file is used in the AICPA Guide to Audit Analytics, Exhibit B-3.

The audit task is to use production and weather-related information to explain revenue and later assess whether expected revenue is consistent with the observed data. Start by checking what variables are available and what one monthly record looks like.

The `head` function shows the values of the variables (columns) in the first few observations (rows) from the dataset.

```{r}
head(USSteamCo)
```

Use this output as a data-readiness check: identify the response variable, the candidate explanatory variables, and the monthly structure before fitting any model.

### Exercise 5.3. Summarizing and plotting the data

This exercise explores the data before modelling. The practical question is whether the variables show ranges, patterns, or relationships that make a regression model plausible for an audit analytics procedure.

The `summary` command produces summary statistics of all variables in the dataset, the length (number of data points) for character types (such as `month`) and distribution statistics for numeric types.

```{r}
summary(USSteamCo)
```

Check the output for scale and completeness. Large differences in units are not a problem by themselves, but they matter when you interpret coefficients and plots later.

We create a dummy variable to distinguish between heating months and cooling months. Such a variable would have the value 1 in heating months and 0 in cooling months. The effect of this variable on the trajectory of the regression line is explained in Section 5.4.4.

The vector below contains one January-to-December pattern of 12 observations. Because the dataset contains 48 monthly observations, `R` recycles this 12-month pattern across the four years automatically. Later, when we include `summer` in a regression model, the `lm` function treats the 0/1 values as an indicator variable and estimates how the intercept and slope differ between the two seasons.


```{r}
USSteamCo$summer <- c(0, 0, 0, 0, 1, 1, 1, 1, 1, 0, 0, 0)
```

We reformat the month variable, that has a character type, into the `date` variable.

```{r}
USSteamCo$date <- seq(as.Date("2011-01-01"),
  by = "month", length.out = 48
)
```

We split the data file into an estimation set and a hold-out set.

```{r}
USSteamCoEstim <- USSteamCo[1:36, ]
USSteamCoHold <- USSteamCo[37:48, ]
```

The estimation set is used to build the model; the hold-out set is kept separate so model performance can later be checked on months that were not used for fitting.


Figure 5.2 shows the histograms of the data in the US SteamCo estimation set.

```{r}
hist_revenue <-
  ggplot(USSteamCoEstim, aes(x = revenue)) +
  geom_histogram(binwidth = 4000000, fill = ada_colour("blue")) +
  scale_x_continuous(labels = label_comma())

hist_production <-
  ggplot(USSteamCoEstim, aes(x = production)) +
  geom_histogram(binwidth = 50000, fill = ada_colour("blue")) +
  scale_x_continuous(labels = label_comma()) +
  theme(plot.margin = unit(c(0, 1, 0, 0), "cm")) # top, right, bottom, left

hist_coolDD <-
  ggplot(USSteamCoEstim, aes(x = coolDD)) +
  geom_histogram(binwidth = 50, fill = ada_colour("blue")) +
  scale_x_continuous(labels = label_comma())

hist_heatDD <-
  ggplot(USSteamCoEstim, aes(x = heatDD)) +
  geom_histogram(binwidth = 100, fill = ada_colour("blue")) +
  scale_x_continuous(labels = label_comma())

histograms_plot <- grid.arrange(hist_revenue, hist_production, hist_coolDD, hist_heatDD, ncol = 2)
```

Figure 5.3 shows the time-series plots for the same dataset.

```{r}
# Calculate the range of revenue and production
rev_range <- range(USSteamCoEstim$revenue, na.rm = TRUE)
prod_range <- range(USSteamCoEstim$production, na.rm = TRUE)

# Create the plot
time_series_plot <- ggplot(data = USSteamCoEstim, aes(x = date)) +
  # First time series (revenue)
  geom_line(aes(y = revenue), colour = ada_colour("blue")) +
  # Second time series (production), rescaled to match revenue range
  geom_line(aes(y = scales::rescale(production, to = rev_range)),
    colour = ada_colour("red")
  ) +
  scale_y_continuous(
    name = "Revenue ($)",
    labels = scales::label_comma(),
    sec.axis = sec_axis(
      ~ scales::rescale(., from = rev_range, to = prod_range),
      name = "Production (x 1000 lb)",
      labels = scales::label_comma()
    )
  ) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y.left = element_text(color = ada_colour("blue")),
    axis.title.y.right = element_text(color = ada_colour("red"))
  )
time_series_plot
```


Figure 5.1 shows a scatter plot of the independent variable production ($x$-axis) and the dependent variable revenue ($y$-axis) for the Case: US SteamCo. The plot is an easy means to get a feel for the relationship.

```{r}
scatter_plot <- ggplot(USSteamCoEstim, aes(x = production, y = revenue)) +
  geom_point(color = ada_colour("blue")) +
  scale_x_continuous(name = "Production (x 1000 lb)", labels = scales::label_comma()) +
  scale_y_continuous(name = "Revenue ($)", labels = scales::label_comma())
scatter_plot
```

The enhanced scatter plot in Figure 5.4 is obtained by running the following code.

```{r}
scatterplot(revenue ~ production,
  data = USSteamCoEstim,
  smooth = FALSE
)
```

Use the plots together rather than separately. The histograms show scale and spread, the time-series plot shows whether the estimation period behaves consistently over time, and the scatter plots show whether production is a reasonable first explanatory variable for revenue.

### Exercise 5.4. The base model `mod.0`

Figure 5.4 shows that there is a positive (increasing) relationship: as production increases, so does revenue.

This exercise fits the first regression model for the audit analytics procedure. The goal is not to choose the final model yet, but to establish a simple baseline that links revenue to production.

We use the function `lm` (for linear model) to estimate the coefficient values of $\beta_0$ and $\beta_1$. This is performed on the data of the estimation set `USSteamCoEstim`, as explained in Section 5.1.1. A brief summary of the model is obtained with `car::brief()`, which is part of the `car` package loaded in Exercise 5.1.


```{r}
ussteamco.mod.0 <- lm(revenue ~ production, data = USSteamCoEstim)
car::brief(ussteamco.mod.0)
```
The estimate of the intercept $\beta_0$ is $b_0 = 4,897,663$, and the estimate of the slope $\beta_1$ is $b_1 = 18.99$. In practical terms, this baseline model estimates expected revenue from production alone. Before relying on it, we still need to inspect fit, residuals, and whether other available variables improve the estimate.

### Exercise 5.5. The full summary of `mod.0`

This exercise opens the full model output for the baseline model. The practical task is to identify which parts of the output speak to model fit, coefficient evidence, and unexplained variation.

A full summary of `mod.0` can be obtained with the `summary` function.

```{r}
summary(ussteamco.mod.0)
```
In addition to the statistics produced by `car::brief()`, it provides insight into the residuals (minimum, first quartile, median, third quartile, and maximum), the $t$- and $p$-values of each of the coefficients and their significance, the adjusted $r^2$, and the $F$ statistic with its $p$-value.


Read the output as a diagnostic summary, not as a final audit conclusion. The coefficient table indicates whether production contributes to the model, while residual spread and $r^2$ indicate how much uncertainty remains.

### Exercise 5.6. Correlations and correlogram

This exercise checks pairwise relationships among candidate variables. The practical goal is to see which variables may help explain revenue and whether some predictors move together strongly enough to complicate interpretation.

The correlations in Table 5.3 are calculated using the `cor` function.

```{r}
(cor_ussteam <- cor(USSteamCoEstim[, 2:5]))
```

Use the signs and magnitudes as a screening tool. A strong correlation with revenue suggests potential explanatory value; strong correlations among predictors are a warning to interpret later coefficients with care.

The correlogram in Figure 5.5 is produced by

```{r echo=FALSE}
corrplot(cor_ussteam,
  tl.col = "black",
  col = colorRampPalette(c(ada_colour("red"), "white", ada_colour("blue")))(100),
  col.lim = c(-1, 1)
)
```

### Exercise 5.7. Modelling strategies

This exercise compares automated model-selection strategies. The practical task is to observe how candidate predictors enter or leave the model, then treat the selected model as a proposal that still needs audit judgment and diagnostics.

The `step` function can be used to implement both backward and forward methods. The `direction` argument specifies whether the stepwise algorithm should proceed forward (`forward`), backward (`backward`), or both (`both`). The `scope` argument specifies the range of models to be searched, whereas the `k` argument may be used to use BIC instead of the default AIC as a criterion.


We start with the forward method. From a model that only uses the constant (`revenue $\sim$ 1`), we add variables one by one. The variable that the largest effect on the criterion, in this example, the default AIC, is selected for inclusion in the model.

```{r}
model_forward <- lm(revenue ~ 1, data = USSteamCoEstim)
model_forward <- step(model_forward,
  direction = "forward",
  scope = formula(~ production + coolDD + heatDD)
)
```
Starting with AIC = 1114.66 of a model with only a constant, `R` evaluates the inclusion of each of the available independent variables and their effect on the resulting AIC. In the first step, it appears that including `heatDD` has the best effect on AIC, as it reduces from 1114.66 to 1096.56.

Using this value of AIC = 1096.56 as the new starting value, `R` evaluates if inclusion of any of the remaining variables has a favorable effect. It finds that including `production` can further reduce the AIC to 1046.11.


In the last step, we see that including `coolDD` does not further reduce the AIC, leading to the final model `revenue $\sim$ heatDD + production`. The summary of this model is:

```{r}
summary(model_forward)
```

At this point, record which variables were selected and which were excluded. The learning point is that AIC-driven selection helps organize candidate models, but it does not replace checking whether the chosen predictors make sense for the audit setting.

Next, we show an example of the backward method, starting with the full model and eliminating variables one by one.

```{r}
model_backward <- lm(revenue ~ production + coolDD + heatDD,
  data = USSteamCoEstim
)
model_backward <- step(model_backward,
  direction = "backward"
)
summary(model_backward)
```

We observe that this strategy results into the exact same model.

To run the stepwise procedure in both directions, we use the following code.

```{r}
fit_both <- lm(revenue ~ 1, data = USSteamCoEstim)
fit_both <- step(fit_both,
  direction = "both",
  scope = formula(~ production + coolDD + heatDD)
)
summary(fit_both)
```

When forward, backward, and both-direction selection end at the same formula, the selection result is more stable. That agreement is useful evidence for model-building, but the model still needs residual and influence checks before it can support an audit expectation.

### Exercise 5.8. Multiple independent variables `mod.1`

This exercise fits a model with multiple independent variables. The practical question is whether adding weather-related predictors improves the expected-revenue model enough to justify the extra complexity.

We now regress $y$ = `revenue` on $x_1$ = `production`, $x_2$ = `coolDD`, and $x_3$ = `heatDD`.

```{r}
ussteamco.mod.1 <- lm(revenue ~ production + coolDD + heatDD,
  data = USSteamCoEstim
)
summary(ussteamco.mod.1)
```

Adding the additional variables `coolDD` and `heatDD` has improved the fit of the model. The residual standard deviation decreased from 4,112,487 for `mod.0` to 1,980,698 for `mod.1`. Similarly, $r^2$ improved from 0.397 to 0.868. These metrics suggest that the inclusion of relevant variables enhances the explanatory power of the model, demonstrating the importance of carefully chosen criteria in stepwise selection methods.

For audit use, focus on what changed from `mod.0`: the unexplained variation is smaller and the model now reflects operational drivers beyond production. The next question is whether the relationship differs by season.

## Exercise 5.9. Interactions

In Section 5.4.4 we developed `mod.2`, that incorporates two separate regression lines. In addition to the three independent variables, we also account for their interactions with the dummy variable `summer`.

This exercise tests whether one revenue relationship is sufficient across the year. The practical task is to allow the slope or intercept to differ between summer and non-summer months, then decide whether that extra structure improves the audit expectation.

The scatterplot in Figure 5.6 is prepared using

```{r}
scat_rev2 <- scatterplot(revenue ~ production | summer,
  data = USSteamCoEstim,
  id = list(n = 8, cex = 0.8),
  regLine = TRUE,
  smooth = FALSE
)
```

```{r}
# Ensure summer is a factor
USSteamCoEstim$summer_fact <- factor(USSteamCoEstim$summer, levels = c(0, 1), labels = c("No", "Yes"))

# Define custom colors
summer_colors <- c("No" = ada_colour("blue"), "Yes" = ada_colour("red"))

# Base plot with manual color scale and separate regression lines
p <- ggplot(USSteamCoEstim, aes(x = production, y = revenue, color = summer_fact)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(values = summer_colors) +
  scale_x_continuous(labels = label_comma()) +
  scale_y_continuous(labels = label_comma()) +
  labs(x = "Production", y = "Revenue", color = "Summer")

p
```


Note the special operator `|`, that can be expressed as 'given the value of'. To call this relationship, we use

```{r}
ussteamco.mod.2 <- lm(
  revenue ~ production * summer +
    coolDD * summer +
    heatDD * summer,
  data = USSteamCoEstim
)
summary(ussteamco.mod.2)
```
The formula specified in the `lm` function, `revenue $\sim$ production * summer + coolDD * summer + heatDD * summer` can also be written as `revenue $\sim$ (production + coolDD + heatDD) * summer`.

As a result, this further improved the fit of the model. The residual standard deviation decreased to 1,680,711, and $r^2$ improved to 0.917.

Interpret this improvement together with the plot. If the seasonal lines describe visibly different revenue patterns and reduce residual spread, the interaction model is a stronger candidate for expected-revenue testing.

## Exercise 5.10. Time lag

This exercise checks whether revenue and production move together at the same time or with a delay. The practical question is whether the model should use current-period predictors only, or whether a lagged relationship deserves attention.

The cross-correlation plot in Figure 5.7 is created by running:

```{r}
with(
  USSteamCoEstim,
  ccf(
    x = production,
    y = revenue,
    lag.max = 4
  )
)
```

Use the largest cross-correlations as a prompt for model judgment. If the strongest relationship occurs at a non-zero lag, a later model may need to reflect timing differences between production activity and recorded revenue.


## Exercise 5.11. Residual plots

This exercise starts the diagnostic review of the selected interaction model. The practical task is to inspect whether residuals look randomly scattered or whether visible patterns suggest a model misspecification.

The residual plots in Figure 5.9 are obtained by:

```{r}
# Add Pearson residuals to the data
USSteamCoEstim$resid <- residuals(ussteamco.mod.2, type = "pearson")

# Extract predictor names
predictors <- all.vars(formula(ussteamco.mod.2))[-1]

# Function to create residual plot
make_resid_plot <- function(data, var) {
  n_unique <- length(unique(data[[var]]))

  p <- ggplot(data, aes(x = .data[[var]], y = resid)) +
    geom_point(color = ada_colour("blue"), alpha = 0.7) +
    geom_hline(yintercept = 0, linetype = "dotted") +
    scale_x_continuous(labels = label_comma()) +
    scale_y_continuous(labels = label_comma()) +
    labs(x = var, y = "")

  if (n_unique > 2) {
    p <- p +
      geom_smooth(
        method = "lm",
        formula = y ~ poly(x, 2),
        se = FALSE,
        color = ada_colour("lightpink")
      )
  } else {
    p <- p +
      geom_smooth(
        method = "lm",
        formula = y ~ x,
        se = FALSE,
        color = ada_colour("lightpink")
      )
  }

  p
}

# Create residual plots for each predictor
plots <- lapply(predictors, function(var) {
  make_resid_plot(USSteamCoEstim, var)
})

# Fitted values vs residuals plot
fitted_data <- data.frame(
  fitted = fitted(ussteamco.mod.2),
  resid = USSteamCoEstim$resid
)

fitted_plot <- ggplot(fitted_data, aes(x = fitted, y = resid)) +
  geom_point(color = ada_colour("blue"), alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_smooth(
    method = "lm",
    formula = y ~ poly(x, 2),
    se = FALSE,
    color = ada_colour("lightpink")
  ) +
  scale_x_continuous(labels = label_comma()) +
  scale_y_continuous(labels = label_comma()) +
  labs(x = "Fitted Values", y = "")

# Top grid: 2x2 layout of the first 4 residual plots
top <- arrangeGrob(
  grobs = plots[1:4],
  ncol = 2
)

# Combine top and bottom
combined <- arrangeGrob(
  top,
  fitted_plot,
  ncol = 1,
  heights = c(2, 1)
)

# Display
grid.newpage()
grid.draw(combined)
```

After creating the plots, look for curves, changing spread, or isolated points. A clean residual pattern supports using the model as an audit expectation; a visible pattern means the model may still be missing an important feature.



## Exercise 5.12. Influence statistics

This exercise checks whether a small number of observations have too much influence on the fitted model. The practical question is whether the audit expectation is robust, or whether particular months drive the model result.

Figure 5.10 provides the Influence index plots. From bottom to top: Leverage points, regression outliers, their respective Bonferroni $p$-values and influential observations.

```{r}
influenceIndexPlot(ussteamco.mod.2)
```

Use the index plots to identify observations that deserve follow-up. An influential month is not automatically an error, but it should be explained before the model is used to support an audit decision.

## Exercise 5.13. Leverage points

This exercise begins a focused diagnostic review of `mod.2`. The practical task is to identify months whose predictor values are unusual enough that they may strongly affect the fitted regression line.

We obtain the hat-values of a model by running the `hatvalues` function on it.

```{r}
hatvals.mod.2 <- hatvalues(ussteamco.mod.2)
```

The three largest hat-values are then summarized by

```{r}
hatvals.mod.2[order(hatvals.mod.2, decreasing = TRUE)[1:3]]
```

Treat these observations as candidates for follow-up. A high leverage value does not mean that the revenue is wrong, but it tells you that the month has unusual explanatory-variable values and can affect the model fit.

The sum of the hat-values is equal to $k + 1$.

```{r}
sum(hatvals.mod.2)
```

This check links the leverage values back to the number of model parameters. It helps verify that the diagnostic is being read on the right scale before moving to outliers and influence.

## Exercise 5.14. Regression outliers

This exercise checks whether any months have unusually large residuals after fitting `mod.2`. The practical question is whether the model misses particular revenue observations badly enough to require explanation.

Figure 5.11 provides the QQ plot for \ttblue{ussteamco.mod.2}, with the three largest Studentized residuals identified.

```{r}
qqPlot(ussteamco.mod.2, id = list(n = 3))
```

Use the labelled observations as a review list. They should be connected back to the business period or underlying records before deciding whether the model is unreliable or whether a particular month is unusual but explainable.

## Exercise 5.15. Influential observations

This exercise combines the previous ideas by measuring how much each observation affects the fitted model. The practical task is to identify months where the audit expectation may depend too heavily on one data point.

We obtain the Cook's distances of a model by running the `cooks.distance` function on it.

```{r}
cooks.mod.2 <- cooks.distance(ussteamco.mod.2)
```

The three largest Cook's distances are then summarized by

```{r}
cooks.mod.2[order(cooks.mod.2, decreasing = TRUE)[1:3]]
```

Compare these observations with the leverage and outlier lists. An observation that appears in several diagnostics deserves priority in the audit follow-up because it may affect both model fit and the resulting expectation.

## Exercise 5.16. Influence plot

Figure 5.12 shows the hat-values, Studentized residuals, and Cook’s distances for `ussteamco.mod.2`. The size of the circles is proportional to the Cook’s distance.

```{r}
influencePlot(ussteamco.mod.2, id = list(n = 3, cex = 0.8))
```

Read the influence plot as a combined diagnostic view. The most important observations are those that are unusual on more than one dimension: high leverage, large residual, or large Cook's distance.

## Exercise 5.17. Winsorizing observation # 22

This exercise performs a sensitivity analysis for observation 22. The goal is not to declare the original value wrong automatically, but to see how much the model changes if the influence of that observation is reduced.

Define the new estimation set. For this sensitivity model, we replace the original value of observation 22 with its winsorized value.

```{r}
USSteamCoEstim2 <- USSteamCoEstim
p5 <- quantile(ussteamco.mod.2$residuals, 0.05)
fit_22 <- fitted(ussteamco.mod.2)[22]
USSteamCoEstim2$revenue[22] <- fit_22 + p5
```

We then fit a new linear model.

```{r}
ussteamco.mod.3 <- lm(revenue ~ (production + heatDD + coolDD) * summer,
  data = USSteamCoEstim2
)

car::brief(ussteamco.mod.3)
```

Compare this brief output with `mod.2`. If the model changes materially, the audit team should understand observation 22 before relying on the regression expectation.

## Exercise 5.18. Multicollinearity

This exercise checks whether predictors overlap so strongly that individual coefficient estimates become hard to interpret. The practical question is whether the model can still support a clear explanation of the revenue expectation.

To assess multicollinearity for a model without interaction terms, such as `mod.1`, we use the standard `vif()` function.

```{r}
vif(ussteamco.mod.1)
```

A common rule of thumb is that VIF values greater than 10 indicate substantial multicollinearity. Since none of the VIF values exceeds this threshold, we conclude that there is no evidence of serious variance inflation in `mod.1`.

For models that include interaction terms, such as `mod.3`, we use the `vif()` function with the additional option `type = "predictor"`. This reports the Generalized Variance Inflation Factor (GVIF) and its adjusted value, \(GVIF^{1/(2df)}\).

```{r}
vif(ussteamco.mod.3, type = "predictor")
```

The adjusted GVIF places predictors with different degrees of freedom on a comparable scale. For a predictor with one degree of freedom, the adjusted GVIF equals the square root of the ordinary VIF. Consequently, the conventional VIF threshold of 10 corresponds to an adjusted GVIF of approximately \(\sqrt{10} \approx 3.16\). Since none of the reported values exceeds 3.16, we conclude that `mod.3` does not exhibit problematic multicollinearity.

For audit interpretation, this means that the selected predictors do not appear to be so redundant that the model explanation becomes unstable.

## Exercise 5.19. Variance and Total Sum of Squares

This exercise connects the variance of revenue to the total variation that regression models try to explain. The practical task is to understand the denominator behind later measures of explained and unexplained variation.

Table 5.5 provides the analysis of variance for `mod.0`. We apply the `anova()` command to perform this analysis.

```{r}
anova(ussteamco.mod.0)
```

The variance of $y$, $s_y^2 = \frac{\sum(y_i - \bar{y})^2}{n - 1}$, is obtained from

```{r}
var(USSteamCoEstim$revenue)
```

Total Sum of Squares $\sum(y_i - \bar{y})^2$ is calculated as

```{r}
mean_y <- mean(USSteamCoEstim$revenue)
variation <- USSteamCoEstim$revenue - mean_y
squared_variation <- variation^2
(total_sum_of_squares <- sum(squared_variation))
```

Total Sum of Squares is the sum of the elements in the `Sum Sq` column of the Analysis of Variance Table.

```{r}
anova.mod.0 <- anova(ussteamco.mod.0)
anova.mod.0$`Sum Sq`[1] + anova.mod.0$`Sum Sq`[2]
```

When we divide Total Sum of Squares by $n - 1$, we get the variance of $y$.

```{r}
total_sum_of_squares / 35
```

This shows how variance is linked to variation.

Keep this calculation in mind when reading model-comparison tables: better models reduce unexplained variation, but the benchmark is the total variation in the response.

## Exercise 5.20. Anova Type I and Type II

This exercise compares two ways to attribute explained variation to predictors. The practical reason is that variable importance can look different when the calculation depends on the order in which predictors enter the model.

We now turn to a model with multiple independent variables, for example, `mod.1`, introduced in Section 5.4.3.

```{r}
(anova.mod.1 <- anova(ussteamco.mod.1))
```

Declaring the independent variables in a different order shows that the calculation of the sum of squares is in sequential order.

```{r}
ussteamco.mod.1b <- lm(revenue ~ heatDD + coolDD + production,
  data = USSteamCoEstim
)
(anova.mod.1b <- anova(ussteamco.mod.1b))
```
To address this issue, an alternative version of the analysis of variance, referred to as Type II tests, is used. Instead of measuring the contribution of each independent variable sequentially, the analysis shows the contribution of each independent variable if it was the last independent variable added to the model. The analysis is obtained using the command `Anova` (with capital A) from the `car` package.

```{r}
Anova(ussteamco.mod.1)
```

We see that the marginal effect of adding `production`, assuming `coolDD` and `heatDD` have already been entered into the model, is 1.8260e+14. This is equal to the result obtained from the earlier `anova` command on `mod.1b`. We also see that the added value of entering `coolDD` into the model is limited: its sum of squares is much lower than that of the other two variables, and its related $F$ value is not significant.

We have no clear preference for either `anova` or `Anova`, they both have their merits. `anova` decomposes TSS, but that decomposition is dependent on the order in which variables are entered into the equation. `Anova` shows the marginal effect of each independent variable, but these marginal effects do not add to TSS.

For the workshop, use this as a caution against over-interpreting a single variable-importance table. Ask whether the result answers the audit question: which variables add useful explanatory value after accounting for the others?

## Exercise 5.21. Anova with interactions

This exercise applies the same analysis-of-variance idea to the sensitivity model with interactions. The practical task is to see which terms still explain variation once seasonal relationships are included.

The analysis of variance of `mod.3` is as follows:

```{r}
(anova.mod.3 <- anova(ussteamco.mod.3))
```

Use this output as a model-review aid, not as the only selection rule. Interaction models should remain interpretable and defensible for the audit expectation.

## Exercise 5.22. Comparing models

This exercise brings the candidate models together. The practical task is to compare fit measures only when the models are based on comparable response data.

We compare the performance of models with four different statistics, the Multiple $r^2$, Adjusted $r^2$ (both obtained from the `summary` function), and the AIC and BIC values. These last two can be called with their own respective commands.

```{r}
AIC(ussteamco.mod.0)
AIC(model_backward)
AIC(ussteamco.mod.1)
AIC(ussteamco.mod.2)
```
The AIC values shown above can be compared directly because all four models are fitted to the same response data (`USSteamCoEstim`). Model `mod.3` is fitted after winsorizing observation 22 and therefore uses a modified response variable. Since the likelihood is computed from the observed responses, the AIC (and BIC) of `mod.3` is not directly comparable with those of the earlier models. Instead, `mod.3` should be regarded as a sensitivity analysis demonstrating the effect of mitigating the influence of an outlying observation.

The AIC value of the `model_backward` does not quite match the output of the `step` function. This can be explained as follows.

The Akaike Information Criterion (AIC) provides a penalized measure of model fit, defined as
$\text{AIC} = -2 \log \hat{L} + 2k$,
where $\log \hat{L}$ is the maximized log-likelihood of the model and $k$ is the number of estimated parameters.

For a linear regression model with normally distributed errors, this log-likelihood takes the form
$\log \hat{L} = -\frac{n}{2} \left[ \log(2\pi) + 1 + \log\left(\frac{\text{SSE}}{n}\right) \right]$,
leading to the AIC expression
$\text{AIC} = n \left[\log(2\pi) + 1 + \log\left(\frac{\text{SSE}}{n}\right)\right] + 2k$.
In contrast, the `step()` function in R reports AIC values based on a deviance-like approximation, that omits constant terms unrelated to model comparison:
$\text{AIC}_{\text{step}} = n \log\left(\frac{\text{SSE}}{n}\right) + 2k$.
The difference between the two expressions is a constant shift of
$n \cdot (\log(2\pi) + 1) + 2$.
The term $n \cdot (\log(2\pi) + 1)$ reflects the part of the log-likelihood that depends only on the sample size and the assumption of normally distributed errors.
The additional +2 arises because R’s AIC() function evaluates the likelihood based on a model with an estimated variance parameter $\sigma^2$, introducing one more estimated parameter than is typically counted in the model formula (e.g., intercept and slopes). This shifts the effective count of parameters k by 1 and contributes an extra penalty of 2 in the AIC expression.


Similarly, the BIC values are obtained as follows.
```{r}
BIC(ussteamco.mod.0)
BIC(model_backward)
BIC(ussteamco.mod.1)
BIC(ussteamco.mod.2)
```

The same consideration applies to the BIC. The BIC values of `mod.0`, `model_backward`, `mod.1`, and `mod.2` may be compared directly because they are based on the same response data. The BIC of `mod.3` should not be interpreted as evidence that the winsorized model is better or worse than the previous models, because it is based on a different dataset.

Before moving to assumption tests, state which model is currently the best candidate and why. Separate fit, interpretability, and sensitivity to observation 22 in that explanation.

## Exercise 5.23. Testing normality

This exercise starts the formal assumption checks for `mod.3`. The practical question is whether the residual distribution is consistent enough with the modelling assumptions to support inference from the model.

We can assess normality issues with a histogram or with a QQ plot. The histogram in Figure 5.14 is obtained by:

```{r}
# Calculate normalized residuals
nres <- ussteamco.mod.3$residuals / sd(ussteamco.mod.3$residuals)
model3 <- as.data.frame(nres)
hist_stand_res <-
  ggplot(model3, aes(nres)) +
  geom_histogram(aes(y = after_stat(density)),
    breaks = seq(-2.55, 1.65, by = 0.7),
    fill = ada_colour("blue")
  ) +
  stat_function(fun = dnorm, args = list(
    mean = mean(model3$nres),
    sd = sd(model3$nres)
  ))
hist_stand_res
```

The QQ plot is produced by:

```{r}
qqPlot(ussteamco.mod.3)
```


The Shapiro-Wilk normality test is executed through the `shapiro.test` function from the `stats` package.

```{r}
shapiro.test(ussteamco.mod.3$residuals)
```

Interpret the visual checks and the Shapiro-Wilk test together. A small p-value would suggest that the residuals depart from normality enough to question normal-theory intervals or tests.

## Exercise 5.24. Testing heteroskedasticity

This exercise checks whether residual variability changes across fitted values or predictors. The practical question is whether prediction uncertainty is reasonably stable across the range of the model.

To test the assumption of constant variance, we can employ visual means, such as the `residualPlots` in Figure 5.15. We can construct it as follows.

```{r}
residualPlots(ussteamco.mod.3,
  layout = c(2, 3),
  quadratic = FALSE, linear = TRUE
)
```
The formal statistical test is the Breusch-Pagan test.

```{r}
bptest(ussteamco.mod.3)
```

Use the plot to look for changing spread and the Breusch-Pagan test as formal support. Evidence of heteroskedasticity would make standard errors and prediction intervals less reliable without adjustment.

## Exercise 5.25. Testing autocorrelation

This exercise checks whether residuals are related over time. Because the SteamCo data are monthly, the practical question is whether one month's unexplained revenue helps predict the next month's unexplained revenue.

We use the `pacf` plot from Figure 5.17 to see if autocorrelation is present, and if so, which order is significant.

```{r}
pacf(ussteamco.mod.3$residuals)
```

The Breusch-Godfrey test is executed through `bgtest`.

```{r}
bgtest(ussteamco.mod.3, order = 3, type = "Chisq")
```

The $p$-value of the test statistic is 0.06737. Therefore, we reject the null hypothesis at levels below 0.06737, suggesting marginal evidence of autocorrelation in the residuals.
Note that the value labelled \texttt{LM test} in the software output is the Breusch--Godfrey test statistic.

For audit use, treat this as a warning rather than a final rejection of the model. Marginal autocorrelation suggests that timing effects may still be present, so the next exercise tests a correction.

## Exercise 5.26. First-order autocorrelation correction

A practical way to address first-order autocorrelation is to apply an AR(1) transformation. This method estimates the autocorrelation parameter and transforms the regression model by subtracting a multiple of the lagged value from the dependent variable and from each predictor. The transformed model can then be estimated by ordinary least squares.

We first estimate an AR(1) model on the residuals from `mod.3`.

```{r}
(rho <- arima(ussteamco.mod.3$residuals,
  order = c(1, 0, 0)
)$coef[1])
```

Read \(\rho\) as the estimated strength of first-order residual dependence. A value near zero would mean little adjustment; a larger value means the transformed model will more strongly account for timing.

Then we transform all variables using the estimated \(\rho\). When the model includes interaction terms, we first construct the interaction terms in their original form and then transform those interaction terms. We should not multiply separately transformed variables.

```{r}
# Create lagged variables, dropping the first observation
n <- nrow(USSteamCoEstim2)

# Current period t and lagged period t - 1
trans <- USSteamCoEstim2[-1, ]
lag <- USSteamCoEstim2[-n, ]

# Construct original interaction terms first
trans$prod_summer <- trans$production * trans$summer
trans$heat_summer <- trans$heatDD * trans$summer
trans$cool_summer <- trans$coolDD * trans$summer

lag$prod_summer <- lag$production * lag$summer
lag$heat_summer <- lag$heatDD * lag$summer
lag$cool_summer <- lag$coolDD * lag$summer

# Transform dependent variable
trans$revenue_adj <- trans$revenue - rho * lag$revenue

# Transform independent variables
trans$production_adj <- trans$production - rho * lag$production
trans$heatDD_adj <- trans$heatDD - rho * lag$heatDD
trans$coolDD_adj <- trans$coolDD - rho * lag$coolDD
trans$summer_adj <- trans$summer - rho * lag$summer

# Transform interaction terms
trans$prod_summer_adj <- trans$prod_summer - rho * lag$prod_summer
trans$heat_summer_adj <- trans$heat_summer - rho * lag$heat_summer
trans$cool_summer_adj <- trans$cool_summer - rho * lag$cool_summer
```

We fit a new model using the transformed variables.

```{r}
ussteamco.mod.4 <- lm(
  revenue_adj ~
    production_adj + heatDD_adj + coolDD_adj +
    summer_adj + prod_summer_adj +
    heat_summer_adj + cool_summer_adj,
  data = trans
)

summary(ussteamco.mod.4)
```

This transformation follows $y_t - \rho y_{t-1} = \beta_0(1-\rho) + \sum_j \beta_j(x_{j,t} - \rho x_{j,t-1}) + u_t$.

For interaction terms, the interaction is treated as a predictor in its own right. Thus, if \(z_t=x_t s_t\), the transformed interaction is $z_t-\rho z_{t-1} = x_t s_t-\rho x_{t-1}s_{t-1}$.

It is not obtained by multiplying the transformed variables. The first observation is discarded because no lagged value is available. After fitting the transformed model, we check whether autocorrelation is reduced by observing the spikes in the `pacf` plot and running the Breusch-Godfrey test.

```{r}
pacf(ussteamco.mod.4$residuals)

bgtest(ussteamco.mod.4, order = 3)
```

Compare the PACF plot and Breusch-Godfrey result with Exercise 5.25. The learning point is whether the correction reduced the autocorrelation problem enough for the model to remain a usable audit expectation.
## Exercise 5.27. Constructing a refined model

This exercise constructs the final candidate model for the hold-out expectation. The practical task is to simplify the interaction model while preserving a structure that remains interpretable for audit use.

When a model includes interaction terms, it is important to respect the hierarchical structure—interaction terms should not be included without their corresponding main effects. The base `R` function `step()` enforces this principle automatically by removing interaction terms before dropping associated main effects. Preserving model hierarchy is essential for ensuring interpretability and theoretical coherence, especially in models intended for inference.


```{r}
ussteamco.mod.5 <- lm(revenue ~ (production + coolDD + heatDD) * summer,
  data = USSteamCoEstim2
)
ussteamco.mod.5 <- step(ussteamco.mod.5,
  direction = "backward"
)
```

Treat the refined model as a candidate, not as an automatic final answer. The next exercises check whether its coefficients, assumptions, and overall significance support using it to create expectations.

## Exercise 5.28. Coefficient test of the refined model

This exercise inspects the coefficient-level evidence for the refined model. The practical question is which predictors still make a meaningful contribution after the model has been simplified.

The summary of this model is:

```{r}
options(scipen = 999)
options(digits = 9)
summary(ussteamco.mod.5)
```

Use the coefficient table to identify the terms that drive the expectation. Terms with weak evidence should be interpreted cautiously, especially when they affect the audit expectation for the hold-out period.

## Exercise 5.29. Re-running the model assumption tests

We re-run the model assumption tests in the same order used earlier in the chapter: normality first, then heteroskedasticity, and finally autocorrelation.

```{r}
shapiro.test(ussteamco.mod.5$residuals)
bptest(ussteamco.mod.5)
bgtest(ussteamco.mod.5, order = 3, type = "Chisq")
```

Read these tests as a final reliability screen. If the refined model still shows serious assumption problems, its predictions may need adjustment, corroboration, or a different modelling approach before they support an audit conclusion.

## Exercise 5.30. Testing significance

The $F$ statistic used for testing the significance of the model as a whole can also be obtained from the \ttblue{summary()} table.


Exercise 5.28 already displayed the full summary table. Here we isolate the $F$ statistic for the refined model so that the model-level significance test is distinguished from the coefficient-level tests.

```{r}
summary(ussteamco.mod.5)
```

The model-level test asks whether the refined model explains revenue better than a model with no predictors. Combine this result with the diagnostics above before carrying `mod.5` forward to the hold-out period.

### Exercise 5.31. Showcase: Visualizing the uncertainty in the model parameters

This exercise is a showcase rather than part of the US SteamCo case workflow. It supports Figure 5.20 by illustrating a general idea: repeated samples from the same underlying regression model produce different fitted lines, even when the true relationship is unchanged.

First, we set the values of the parameters. Play around with these to see the effect on the ultimate graph!

```{r}
set.seed(123) # Set seed for reproducibility
beta_0 <- 2 # True intercept
beta_1 <- 1.5 # True slope
n <- 11 # Number of points to sample
x_min <- 0 # Minimum value of x
x_max <- 10 # Maximum value of x
x_values <- seq(x_min, x_max, length.out = n) # x values modelled
sigma <- 5 # Standard deviation of the noise
rep <- 200 # Number of replications
alpha <- 0.05 # Significance level for 95% confidence interval
x_extended_min <- x_min - (x_max - x_min) / 2 # Extended x range minimum
x_extended_max <- x_max + (x_max - x_min) / 2 # Extended x range maximum
x_extended <- seq(x_extended_min, x_extended_max, length.out = 100) # Plot range
```

We first calculate the true regression line, along with its confidence interval.

```{r}
# Create a data frame for the true line over the extended range
true_line <- data.frame(x = x_extended, y = beta_0 + beta_1 * x_extended)

# Calculate the standard error of the true regression line
mean_x <- mean(x_values) # Mean of the x values
S_xx <- sum((x_values - mean_x)^2) # Sum of squares of x deviations
se_fit_true <- sigma * sqrt(1 / n + (x_extended - mean_x)^2 / S_xx) # Standard error formula for regression

# Calculate the theoretical confidence intervals based on the true model
t_value <- qt(1 - alpha / 2, df = n - 2) # Critical t-value for 95% CI
ci_upper_true <- (beta_0 + beta_1 * x_extended) + t_value * se_fit_true # Upper bound
ci_lower_true <- (beta_0 + beta_1 * x_extended) - t_value * se_fit_true # Lower bound

# Create a data frame for the confidence interval bounds based on the true regression line
ci_bounds_true <- data.frame(x = x_extended, ci_lower = ci_lower_true, ci_upper = ci_upper_true)
```

Next, we sample $n$ points from a normal distribution, to generate $y$ values for each $x$ value and fit the corresponding regression line. We repeat this process `rep` times.

```{r}
# Initialize a data frame to store the simulated regression lines
simulated_lines <- data.frame()

# Repeat the process `rep` times to simulate multiple regression lines
for (i in 1:rep) {
  # Sample n points from the relationship y = beta_0 + beta_1 * x + epsilon
  epsilon <- rnorm(n, mean = 0, sd = sigma) # Random noise
  y_values <- beta_0 + beta_1 * x_values + epsilon

  # Fit a linear model
  model <- lm(y_values ~ x_values)

  # Predict fitted values for the extended range of x
  predicted_values <- predict(model, newdata = data.frame(x_values = x_extended))

  # Store the fitted line over the extended x range
  simulated_lines <- rbind(
    simulated_lines,
    data.frame(
      x = x_extended,
      y = predicted_values,
      rep = i
    )
  )
}
```

All the elements created earlier are now plotted on a single canvas.

```{r}
regression_lines_plot <- ggplot() +
  geom_line(
    data = simulated_lines, aes(x = x, y = y, group = rep),
    color = ada_colour("blue"), alpha = 0.1
  ) + # Plot the simulated regression lines
  geom_line(data = true_line, aes(x = x, y = y), color = ada_colour("lightpink"), linewidth = .5) + # Plot the true line
  geom_line(
    data = ci_bounds_true, aes(x = x, y = ci_lower),
    color = ada_colour("lightpink"), linewidth = .5, linetype = "solid"
  ) + # Lower bound (true line)
  geom_line(
    data = ci_bounds_true, aes(x = x, y = ci_upper),
    color = ada_colour("lightpink"), linewidth = .5, linetype = "solid"
  ) + # Upper bound (true line)
  labs(
    title = paste(rep, " Regression Lines with Confidence Interval Based on True Model"),
    x = "x", y = "y"
  ) +
  ggplot2::theme_bw() +
  xlim(x_extended_min, x_extended_max) # Extend the x-axis limits
regression_lines_plot
```

The resulting graph is shown in Figure 5.20.

Use the figure as a reminder that a fitted regression line is an estimate. The same modelling process can produce different lines when the sample changes, which is why prediction intervals matter in the audit expectation exercises.

## Exercise 5.32. Showcase: Confidence and prediction intervals

This exercise is also an illustration rather than a case step. It stays in the student notebook because it introduces the visual distinction between a confidence interval for the mean response and a prediction interval for a new observation before we return to the case-based forecasting exercises.

Figure 5.18 shows a regression line with a dotted confidence interval and a shaded prediction interval. It is created with the following code.

```{r}
# Calculate the standard error for prediction interval
alpha_pred <- 0.01 # For 99% prediction interval
t_value_pred <- qt(1 - alpha_pred / 2, df = n - 2)
se_pred <- sigma * sqrt(1 + 1 / n + (x_extended - mean_x)^2 / S_xx)

# Calculate 99% prediction interval bounds
pi_upper_true <- (beta_0 + beta_1 * x_extended) + t_value_pred * se_pred
pi_lower_true <- (beta_0 + beta_1 * x_extended) - t_value_pred * se_pred

pi_bounds_true <- data.frame(
  x = x_extended,
  pi_lower = pi_lower_true,
  pi_upper = pi_upper_true
)

intervals_plot <- ggplot() +
  geom_ribbon(
    data = pi_bounds_true,
    aes(x = x, ymin = pi_lower, ymax = pi_upper),
    fill = ada_colour("blue"), alpha = 0.15
  ) + # 99% prediction interval as a ribbon
  geom_line(
    data = ci_bounds_true,
    aes(x = x, y = ci_lower),
    color = ada_colour("blue"), linewidth = .5, linetype = "dashed"
  ) + # 95% CI lower
  geom_line(
    data = ci_bounds_true,
    aes(x = x, y = ci_upper),
    color = ada_colour("blue"), linewidth = .5, linetype = "dashed"
  ) + # 95% CI upper
  geom_line(
    data = true_line,
    aes(x = x, y = y),
    color = ada_colour("blue"), linewidth = 1
  ) + # True regression line
  labs(
    x = "x", y = "y"
  ) +
  xlim(x_extended_min, x_extended_max)
intervals_plot
```

The shaded prediction interval is wider because it includes uncertainty for a new observation, not only uncertainty in the average response. That distinction is the key bridge back to the hold-out revenue predictions in the next exercise.


## Exercise 5.33. Creating expectations for the hold-out period

This exercise applies the refined model to the 2014 hold-out months. The practical task is to create an expected revenue and prediction interval for each month, then compare those expectations with recorded revenue.

The function `predict` creates expectations, when data from `newdata` are evaluated with a certain `object`. The optional `interval` is used to add bounds for a prediction interval, or, alternatively, a confidence interval. The `level` specifies the confidence level of the interval.



```{r echo=FALSE}
(predictions <- predict(
  object = ussteamco.mod.5,
  newdata = USSteamCoHold,
  interval = "prediction",
  level = 0.99
))
```
Compare the recorded revenue with the expectations.

```{r}
comparison <- cbind(
  Month = rownames(USSteamCoHold),
  Recorded = USSteamCoHold$revenue,
  Lower = round(predictions[, "lwr"], 0),
  Expected = round(predictions[, "fit"], 0),
  Upper = round(predictions[, "upr"], 0)
)

comparison <- as.data.frame(comparison)

comparison$Recorded <- as.numeric(comparison$Recorded)
comparison$Expected <- as.numeric(comparison$Expected)
comparison$Lower <- as.numeric(comparison$Lower)
comparison$Upper <- as.numeric(comparison$Upper)

# Difference = Recorded - Expected
comparison$Difference <- comparison$Recorded - comparison$Expected

comparison
sum(comparison$Recorded)
sum(comparison$Expected)
sum(comparison$Difference)
```

Use the monthly differences as a first exception scan. A month outside the prediction interval or with a large recorded-minus-expected difference should be explained before the annual result is interpreted.


## Exercise 5.34. Plotting expectations against recorded values

This exercise visualizes the same monthly comparison. The practical question is whether any recorded revenue pattern falls outside the model-based expectation rather than only whether individual numbers differ.

In Figure 5.19 we have plotted expectations against the recorded values of revenue, indicating which individual observations fall outside of the 99% prediction interval. The following code is used to create it.

```{r}
# Convert predictions to a data frame and add the date column
pred_df <- data.frame(
  date = USSteamCoHold$date,
  recorded = USSteamCoHold$revenue,
  lwr = predictions[, "lwr"],
  fit = predictions[, "fit"],
  upr = predictions[, "upr"]
)
pred_df$outside <- with(pred_df, recorded < lwr | recorded > upr)

expectations_plot <- ggplot(pred_df, aes(x = date)) +
  # Plot the predicted values as a line
  geom_line(aes(y = fit, color = "Expectation"), linewidth = 1) +
  # Add the prediction interval as a shaded area
  geom_ribbon(aes(ymin = lwr, ymax = upr), fill = ada_colour("blue"), alpha = 0.2) +
  # Plot the actual values, with conditional coloring for outside interval points
  geom_line(aes(y = recorded, color = "Recorded"), linewidth = 1) +
  # Highlight points outside interval in red
  geom_point(data = subset(pred_df, outside), aes(y = recorded), color = ada_colour("lightpink"), size = 2) +
  # Define color for predicted and actual values
  scale_color_manual(
    values = c("Expectation" = ada_colour("blue"), "Recorded" = ada_colour("lightpink")),
    name = ""
  ) +
  labs(
    x = "Month in 2014",
    y = "Revenue ($)"
  ) +
  scale_x_date(
    date_breaks = "1 month", # Adjusts to one gridline per month
    date_labels = "%b" # Optional: shows abbreviated month names
  ) +
  scale_y_continuous(labels = label_comma()) +
  theme(legend.position = "bottom")

expectations_plot
```

Use the plot to identify months that deserve audit follow-up. A point outside the interval is not automatically a misstatement, but it should be reconciled with corroborating evidence or business events.

## Exercise 5.35. Combining 12 monthly predictions

We use Equations 5.29 and 5.30 to construct a prediction interval for total revenue for the year under audit. The annual expectation is the sum of the 12 monthly expectations.

To keep the derivation transparent, we first collect the monthly expectations, then combine them into an annual expectation, and finally compute the prediction standard error of that annual total. The design-matrix calculation accounts for the covariance between the monthly expectations caused by using the same estimated regression coefficients in all 12 months.


```{r}
# Residual degrees of freedom
df_res <- ussteamco.mod.5$df.residual

# Monthly expectations from Exercise 5.34
monthly_expectations <- pred_df$fit

monthly_expectation_table <- data.frame(
  month = format(pred_df$date, "%B"),
  expected_revenue = monthly_expectations
)

monthly_expectation_table
```

The annual expectation is obtained by summing the 12 monthly expectations. This keeps the annual calculation tied directly to the month-by-month audit expectation inspected in the previous exercise.

```{r}
# Combined annual expectation
annual_prediction <- sum(monthly_expectations)
annual_prediction
```

The next step prepares the model information needed to quantify uncertainty in that annual expectation. The design matrix contains one row for each hold-out month and one column for each coefficient in the final regression model.

```{r}
# Construct the design matrix for the hold-out observations
X_hold <- model.matrix(
  formula(ussteamco.mod.5),
  data = USSteamCoHold
)

# Vector of ones for summing the 12 monthly predictions
one <- rep(1, nrow(X_hold))

data.frame(
  quantity = c(
    "hold-out months",
    "model coefficients"
  ),
  value = c(
    nrow(X_hold),
    ncol(X_hold)
  )
)
```

The annual prediction variance has two parts. The first part comes from uncertainty in the estimated regression coefficients. The second part comes from the future residual variation of the 12 monthly observations.

```{r}
# Variance contribution from uncertainty in the estimated regression coefficients
var_mean_annual <- as.numeric(
  t(one) %*% X_hold %*% vcov(ussteamco.mod.5) %*% t(X_hold) %*% one
)

# Variance contribution from the future residual variation of 12 monthly observations
sigma2 <- summary(ussteamco.mod.5)$sigma^2
var_future_annual <- nrow(X_hold) * sigma2

# Total prediction variance of the annual amount
annual_var <- var_mean_annual + var_future_annual

# Prediction standard error of the annual amount
annual_se <- sqrt(annual_var)

annual_variance_components <- data.frame(
  component = c(
    "coefficient uncertainty",
    "future residual variation",
    "total annual prediction variance",
    "annual prediction standard error"
  ),
  value = c(
    var_mean_annual,
    var_future_annual,
    annual_var,
    annual_se
  )
)

annual_variance_components
```

The coefficient-uncertainty component reflects that the regression coefficients are estimated from a sample. The future-residual component reflects that even a well-specified model will not predict each future month exactly. The latter is calculated under the assumption that the future residual errors are uncorrelated.

Finally, we use the residual degrees of freedom to obtain the critical \(t\)-value and construct the 99% prediction interval for annual revenue.

```{r}

# Critical value for a 99% prediction interval
t_score <- qt(0.995, df = df_res)
annual_interval_half_width <- t_score * annual_se

# 99% prediction interval for the annual total
annual_lower <- annual_prediction - annual_interval_half_width
annual_upper <- annual_prediction + annual_interval_half_width

annual_prediction_interval <- data.frame(
  quantity = c(
    "lower bound",
    "annual expectation",
    "upper bound",
    "annual prediction standard error"
  ),
  value = round(
    c(
      annual_lower,
      annual_prediction,
      annual_upper,
      annual_se
    ),
    0
  )
)

annual_prediction_interval
```

Read the annual interval as the range of total recorded revenue that would be expected from the model before considering specific corroborating evidence.

Adjust the March prediction based on the corroborated evidence for the winter conditions and disruptions. This is an audit judgement step: the adjustment should be supported by documented evidence, not used merely to make the model fit the recorded amount.

```{r}
storm_adjustment <- 8000000

pred_df$fit_adj <- pred_df$fit
pred_df$lwr_adj <- pred_df$lwr
pred_df$upr_adj <- pred_df$upr

march_row <- format(pred_df$date, "%b") == "Mar"

pred_df$fit_adj[march_row] <- pred_df$fit[march_row] - storm_adjustment
pred_df$lwr_adj[march_row] <- pred_df$lwr[march_row] - storm_adjustment
pred_df$upr_adj[march_row] <- pred_df$upr[march_row] - storm_adjustment

pred_df$outside_adj <- with(pred_df, recorded < lwr_adj | recorded > upr_adj)
pred_df
```

For the annual calculations in the next exercise, the adjusted annual expectation is the sum of the adjusted monthly expectations.

```{r}
annual_prediction_adj <- sum(pred_df$fit_adj)
annual_prediction_adj
```

The adjusted annual expectation carries the documented March explanation into the annual evaluation. The next exercise compares recorded revenue with this adjusted expectation.

## Exercise 5.36. Level of assurance for the annual prediction

This exercise turns the annual prediction into an audit decision check. The practical task is to compare recorded annual revenue with the adjusted expectation and decide whether the difference is acceptable under the chosen materiality and risk parameters.

We now compare the recorded annual revenue with the annual expectation. Consistent with the convention used throughout the book, the difference is defined as the recorded amount minus the expectation.

We use the numbers adjusted for the winter storm conditions.
```{r}
monthly_expectations_adj <- pred_df$fit_adj

adjusted_monthly_expectation_table <- data.frame(
  month = format(pred_df$date, "%B"),
  adjusted_expected_revenue = monthly_expectations_adj
)

adjusted_monthly_expectation_table
```

The adjusted annual expectation is the sum of these adjusted monthly expectations.

```{r}
annual_prediction <- sum(monthly_expectations_adj)
annual_prediction
```

The recorded annual revenue is the benchmark amount being audited. Subtracting the adjusted expectation gives the observed difference for the year.

```{r}
# Difference convention: recorded amount minus expectation
annual_recorded <- sum(USSteamCoHold$revenue)
annual_difference <- annual_recorded - annual_prediction
annual_difference
```

Standardizing the annual difference by the annual prediction standard error puts the observed difference on the same scale as the decision bounds calculated below.

```{r}
test_statistic <- annual_difference / annual_se

annual_difference_summary <- data.frame(
  quantity = c(
    "recorded annual revenue",
    "adjusted annual expectation",
    "annual difference",
    "annual prediction standard error",
    "standardized difference"
  ),
  value = c(
    annual_recorded,
    annual_prediction,
    annual_difference,
    annual_se,
    test_statistic
  )
)

annual_difference_summary
```

Interpret the sign as well as the size of the difference. A positive difference means recorded revenue is above expectation; a negative difference means recorded revenue is below expectation.

Next, we set the audit parameters and construct the acceptable difference range. The acceptable difference range is expressed around zero, because it is applied to the difference between the recorded amount and the expectation.

```{r}
# Set audit parameters
PM <- 15000000 # Performance materiality
beta <- 0.01 # Efficiency risk beta
confidence_level <- 1 - beta
df <- df_res

# Standardized materiality
delta <- PM / annual_se

# Two-sided critical values
c_L <- qt(beta / 2, df = df)
c_U <- qt(1 - beta / 2, df = df)

# Acceptable difference range for the annual difference
ADR_lower <- c_L * annual_se
ADR_upper <- c_U * annual_se
```

The quantities `c_L` and `c_U` are standardized decision bounds. Multiplying them by `annual_se` converts them into the acceptable difference range for the annual difference.

```{r}
acceptable_difference_range <- data.frame(
  boundary = c(
    "lower acceptable difference",
    "upper acceptable difference"
  ),
  value = c(
    ADR_lower,
    ADR_upper
  )
)

acceptable_difference_range
```

Equivalently, this acceptable difference range can be translated into lower and upper bounds around the annual expectation.

```{r}

# Equivalent acceptable bounds for recorded annual revenue
recorded_lower <- annual_prediction + ADR_lower
recorded_upper <- annual_prediction + ADR_upper

acceptable_recorded_revenue_bounds <- data.frame(
  boundary = c(
    "lower bound",
    "annual expectation",
    "upper bound"
  ),
  value = c(
    recorded_lower,
    annual_prediction,
    recorded_upper
  )
)

acceptable_recorded_revenue_bounds
```

The audit decision compares the observed annual difference with this acceptable range. If the difference falls outside the range, the model points to a potential overstatement or understatement that requires follow-up.

```{r}
annual_decision_inputs <- data.frame(
  quantity = c(
    "observed annual difference",
    "lower acceptable difference",
    "upper acceptable difference"
  ),
  value = c(
    annual_difference,
    ADR_lower,
    ADR_upper
  )
)

annual_decision_inputs
```


The achieved assurance is calculated using the non-central \(t\) distribution. A material overstatement shifts the distribution of the standardized difference to the right. A material understatement shifts it to the left.

```{r}
# Achieved assurance when controlling efficiency risk

assurance_over <- 1 - pt(c_U, df = df, ncp = delta)
assurance_under <- pt(c_L, df = df, ncp = -delta)

achieved_assurance <- data.frame(
  material_difference = c(
    "overstatement",
    "understatement"
  ),
  achieved_assurance_percent = round(
    c(
      assurance_over,
      assurance_under
    ) * 100,
    1
  )
)

achieved_assurance
```

With the current implementation and the audit parameters above, the achieved assurance is approximately 36.9% in each direction.

This assurance level describes how strongly the analytical procedure can detect a material difference under the selected assumptions. It should be considered together with the acceptable-difference conclusion, not read as a stand-alone pass or fail.


Finally, we evaluate the observed annual difference.

```{r}
# Conclusion for the observed annual difference

if (annual_difference < ADR_lower) {
  conclusion <- "Understatement follow-up."
} else if (annual_difference > ADR_upper) {
  conclusion <- "Overstatement follow-up."
} else {
  conclusion <- "Difference acceptable."
}

annual_decision <- data.frame(
  observed_difference = annual_difference,
  lower_acceptable_difference = ADR_lower,
  upper_acceptable_difference = ADR_upper,
  conclusion = conclusion
)

annual_decision
```

Use this conclusion as the documented result of the analytical procedure. If the difference is acceptable, the model does not identify a material unexpected annual revenue difference under the assumptions used here.
