In this section, we will embark on a series of exercises designed to reinforce the concepts covered in the theory sections of this chapter. For these exercises, we will be using \ttblue{Python} as our programming language of choice.

\subsection*{Background on Python}

\ttblue{Python} is a versatile and widely-used programming language known for its simplicity and readability. Developed by Guido van Rossum and first released in 1991, \ttblue{Python} has since gained immense popularity among developers, data scientists, and researchers due to its ease of learning and extensive ecosystem of libraries and tools.

One of the key strengths of \ttblue{Python} is its versatility. From web development to scientific computing, \ttblue{Python} can be used in a wide range of applications. Its clear and concise syntax makes it accessible to beginners, while its powerful features and libraries, such as \ttblue{NumPy}, \ttblue{Pandas}, and \ttblue{Matplotlib}, make it a favorite among experienced programmers for data analysis, machine learning, and more.

\subsection*{Integration with Theory}

All examples and exercises presented in this section are directly linked to the theoretical concepts discussed earlier in the chapter. By applying these concepts in practical scenarios using \ttblue{Python}, you will gain a deeper understanding of their implications and practical utility.

Throughout the exercises, we encourage you to refer back to the corresponding theory sections for clarification and context. This integrated approach ensures a holistic learning experience, where theoretical knowledge is reinforced and applied in real-world contexts.

Let's dive into the exercises and explore the practical applications of the concepts covered in this chapter.

\begin{exercise}
\label{ex:hyper_p}

\textbf{Hypergeometric distribution} \marginnote{Section \ref{sec:hypergeometric_distribution}}

The \ttblue{Python} function \ttblue{hypergeom.pmf} requires four arguments: $k, N, M$, and $n$. The extension \ttblue{pmf} indicates that the function is a probability mass (density) function.

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{KPMG_blue}{hypergeom.pmf(0, 331, 17, 60)}
\textcolor{KPMG_light_blue}{0.03036497699701723}
\textcolor{KPMG_blue}{hypergeom.pmf(1, 331, 17, 60)}
\textcolor{KPMG_light_blue}{0.12145990798806892}
\end{Verbatim}

\emph{Cumulative probabilities} $P(\stoch{k}\leq k)$ (the probability of finding not more than $k$ errors) are obtained with \ttblue{hypergeom.cdf}. For example, if we are interested in the probability of finding no more than one error, we use \ttblue{hypergeom.cdf(1, 331, 17, 60)}. In this case the extension \ttblue{cdf} indicates the cumulative density function.

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{KPMG_blue}{hypergeom.cdf(1, 331, 17, 60)}
\textcolor{KPMG_light_blue}{0.15182488498508614}
\textcolor{KPMG_blue}{hypergeom.pmf(0, 331, 17, 60) + hypergeom.pmf(1, 331, 17, 60)}
\textcolor{KPMG_light_blue}{0.15182488498508614}
\end{Verbatim}

We see that the result for the cumulative function equals that of the sum of the two individual probabilities.

Probabilities in the right-hand tail $P(\stoch{k} > k)$ can be calculated using the \ttblue{hypergeom.sf} function. The extension \ttblue{sf} indicates the survival function.

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{KPMG_blue}{hypergeom.sf(1, 331, 17, 60)}
\textcolor{KPMG_light_blue}{0.8481751150149138}
\end{Verbatim}

\end{exercise}

\begin{exercise}
\label{ex:binom_p}

\textbf{Binomial distribution} \marginnote{Section \ref{sec:binomial_distribution}}

The relevant \ttblue{Python} functions are \ttblue{binom.pmf} for the probabilities $P(\stoch{k} = k)$ and \ttblue{binom.cdf} for cumulative probabilities $P(\stoch{k} \leq k)$. Both functions use the parameters \ttblue{k}, \ttblue{n} and \ttblue{$\pi$ = M / N}.

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{KPMG_blue}{binom.pmf(k = 1, n = 60, p = 17 / 331)}
\textcolor{KPMG_light_blue}{0.1373313791683671}
\textcolor{KPMG_blue}{binom.cdf(1, 60, 17 / 331)}
\textcolor{KPMG_light_blue}{0.17960790177509958}
\end{Verbatim}
\end{exercise}

\begin{exercise}
\label{ex:poiss_p}

\textbf{Poisson distribution} \marginnote{Section \ref{sec:poisson_distribution}}

The functions \ttblue{poisson.pmf} and \ttblue{poisson.cdf} are used for the calculation of probabilities from the Poisson distribution. They require two parameter values as arguments: the number of errors $k$ and the value of $\mu = n\pi$. For the \emph{Case: Number of students} remember that $n = 60$ and $\pi = 17 / 331$.

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{KPMG_blue}{poisson.pmf(k = 1, mu = 60 * 17/ 331)}
\textcolor{KPMG_light_blue}{0.1414043918733124}
\textcolor{KPMG_blue}{poisson.pmf(3, 60 * 17 / 331)}
\textcolor{KPMG_light_blue}{0.22379789843860848}
\end{Verbatim}

The probability of finding one error is 14.14\%.

To calculate the probability of finding no more than one error in the sample of 60 Ph.D.s, use 

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{KPMG_blue}{poisson.cdf(1, 60 * 17 / 331)}
\textcolor{KPMG_light_blue}{0.1872915033537697}
\end{Verbatim}

\end{exercise}

\begin{exercise}
\label{ex:norm_p}

\textbf{Normal distribution} \marginnote{Section \ref{sec:normal_distribution}}

In \ttblue{Python} we use the function \ttblue{norm.cdf}, that uses the arguments \ttblue{x} for the sampling result that we evaluate, \ttblue{loc} for the population mean, and \ttblue{scale} for the standard deviation of the mean.

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{KPMG_blue}{norm.cdf(x = 1012, loc = 1030, scale = 115.26 / math.sqrt(200))}
\textcolor{KPMG_light_blue}{0.013602685762330115}
\end{Verbatim}

Note that the syntax is more elaborate than what we have done so far. In \ttblue{Python} we can choose between passing the values of the arguments only and passing values specifically assigned to the arguments that the function uses.

There are two advantages to using argument names and values. First, the command is easier to interpret by a reviewer. Second, we can pass the arguments in any order. The following commands are therefore equivalent:

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{KPMG_blue}{norm.cdf(1012, 1030, 115.26 / math.sqrt(200))}
\textcolor{KPMG_light_blue}{0.013602685762330115}
\textcolor{KPMG_blue}{pnorm(norm.cdf(x = 1012, loc = 1030, scale = 115.26 / math.sqrt(200))}
\textcolor{KPMG_light_blue}{0.013602685762330115}
\textcolor{KPMG_blue}{norm.cdf(scale = 115.26 / math.sqrt(200), x = 1012, loc = 1030)}
\textcolor{KPMG_light_blue}{0.013602685762330115}
\end{Verbatim}

\end{exercise}

\begin{exercise}
\label{ex:student_p}

\textbf{Student's t distribution} \marginnote{Section \ref{sec:student_t_distribution}}

To calculate the probability found of 14.7\%, we use the function \ttblue{pt.cdf} with arguments \ttblue{x} for the boundary value and \ttblue{df} for the number of degrees of freedom.

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{KPMG_blue}{tval = (1004 - 1030) / (73.78 / math.sqrt(10))}
\textcolor{KPMG_blue}{t.cdf(x = tval, df = 9)}
\textcolor{KPMG_light_blue}{0.14699391173483017}
\end{Verbatim}

The probability is 14.7\%.

\end{exercise}

\begin{exercise}
\label{ex:chisq_p}

\textbf{\texorpdfstring{$\chi^2$}{chi-squared} distribution} \marginnote{Section \ref{sec:chi_squared_distribution}}

The 95\% upper bound on a $\chi^2$ distributed variable with (10 - 1) degrees of freedom is

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{KPMG_blue}{chi2.ppf(q = .95, df = 9)}
\textcolor{KPMG_light_blue}{16.918977604620448}
\end{Verbatim}

Similarly, to calculate the upper bound for $\sigma^2$, we start with the lower bound of a $\chi^2$ distributed variable

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{KPMG_blue}{chi2.ppf(q = .05, df = 9)}
\textcolor{KPMG_light_blue}{3.325112843066815}
\end{Verbatim}

\end{exercise}

\begin{exercise}
\label{ex:fdist_p}

\textbf{F distribution} \marginnote{Section \ref{sec:f_distribution}}

Hypothesis testing is done with the function \ttblue{f.ppf}.

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{KPMG_blue}{f.ppf(q = .95, dfn = 25, dfd = 23)}
\textcolor{KPMG_light_blue}{1.9962706179379208}
\end{Verbatim}

In the example above, the critical value of the $F$-statistic is 1.996. Depending on whether the value of the test statistic is greater or smaller than this value, we then reject the null hypothesis or we do not. 

The probability of obtaining this particular value of the test statistic is calculated with the \ttblue{f.sf} function.

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{KPMG_blue}{f_crit = f.ppf(q = .95, dfn = 25, dfd = 23)}
\textcolor{KPMG_blue}{f.sf(x = f_crit, dfn = 25, dfd = 23)}
\textcolor{KPMG_light_blue}{0.049999999999999975}
\end{Verbatim}

\end{exercise}
