---
title: "Workshop Chapter 4"
author: "Lucas Hoogduin"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---
<!-- SUPPORT-ONLY:START -->
```{r notebook-utils, include=FALSE}
source("../../../R/notebook-utils.R")
```
<!-- SUPPORT-ONLY:END -->

## Hypothesis testing

### Exercise 4.1. Effect of the critical region on the sample size

The results on page 119 are obtained with the `FSaudit` package, first by creating an attribute object, filling it with the parameters of the test, and then applying the `size()` function.
The relevant parameters are the significance level `alpha`, the number of deviations in the population $M$ `popdev`, the population size $N$ `popn`, and the critical region `c`.

<!-- ADA:REQUIRES capability=fsaudit -->
```{r} 
library(FSaudit)
mySample <- att_obj(alpha = .1, popdev = 60, popn = 1200, c = 0)
mySample <- size(mySample)
mySample$n
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.1")
mySample = att_sample(alpha=0.1, popdev=60, popn=1200, c=0).size()
mySample.n
```
<!-- ADA:END -->
The sample size increases as we increase the critical region to `c = 2`.
```{r}
mySample <- size(mySample, c = 2)
mySample$n
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.1")
mySample.size(c=2).n
```
<!-- ADA:END -->

This result is obtained using the default Hypergeometric distribution. To obtain the sample size with the binomial approximation, as in Table 5.3, we specify the distribution when setting up the attribute object.

```{r}
mySample2 <- att_obj(
  alpha = .1,
  tdr = .05,
  c = 2,
  dist = "binom"
)
mySample2 <- size(mySample2)
mySample2$n
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.1")
mySample2 = att_sample(alpha=0.1, tdr=0.05, c=2, dist="binom").size()
mySample2.n
```
<!-- ADA:END -->

### Exercise 4.2. Significance levels

On page 120 we calculated significance levels for the occurrence of finding one or two errors. These probabilities are calculated in `R` with the following code:

<!-- ADA:BEGIN lang=python mode=override kind=narrative -->
On page 120 we calculated significance levels for the occurrence of finding one
or two errors. These probabilities are calculated in Python with the following
code:
<!-- ADA:END -->

```{r}
phyper(q = 1, m = 60, n = 1140, k = 45)
phyper(q = 2, m = 60, n = 1140, k = 102)
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
hypergeom.cdf(1, 1200, 60, 45)
hypergeom.cdf(2, 1200, 60, 102)
```
<!-- ADA:END -->
Refer to Equation 2.1.
Remember that `R` uses `q` for the number of errors found $k$, `m` for the number of deviations in the population $M$, `n` for the number of correct items $N - M$, and `k` for the sample size $n$.

<!-- ADA:BEGIN lang=python mode=override kind=narrative -->
Refer to Equation 2.1. Remember that Python's hypergeometric distribution
function uses the population size $N$, the number of deviations in the
population $M$, the sample size $n$, and the number of errors found $k$.
<!-- ADA:END -->

### Exercise 4.3. Type II error

On page 121 we also calculated the Type II error, for a scenario with no errors allowed in the sample and 24 errors in the population, or a population error rate of $24 / 200 = 0.02$.

```{r}
phyper(0, m = 24, n = 1176, k = 45, lower.tail = FALSE)
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
hypergeom.sf(0, 1200, 24, 45)
```
<!-- ADA:END -->

### Exercise 4.4. One-sided upper bounds

The one-sided upper bounds $p_U$ in Table 5.1 are obtained as follows:

```{r}
upper(popn = 1200, n = 102, k = 0, alpha = 0.10) / 1200
upper(popn = 1200, n = 102, k = 3, alpha = 0.10) / 1200
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
upper_bound(k=0, popn=1200, n=102, alpha=0.10) / 1200
upper_bound(k=3, popn=1200, n=102, alpha=0.10) / 1200
```
<!-- ADA:END -->

### Exercise 4.5. Case: European innovation subsidies
The sample size in the *Case: European innovation subsidies* depends on the critical region chosen. When the null hypothesis $H_0 : M \geq 120,000$ is rejected when the sample yields no errors, the critical region is $\{k | k = 0\}$. We first create an `mus_obj` object, and load it with the parameters.
<!-- ADA:REQUIRES capability=fsaudit -->
```{r}
subsidies <- mus_obj(
  cl = 0.95,
  popBv = 12000000,
  pm = 120000
)
subsidies <- size(subsidies,
  ee = 0
)
subsidies$n
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.5")
subsidies = mus_sample(cl=0.95, popBv=12000000, pm=120000).size(ee=0)
subsidies.n
```
<!-- ADA:END -->
This result is exactly equal to that of the fixed-attribute sample:
```{r}
myAttSample <- att_obj(alpha = 0.05, popn = 12000000, popdev = 120000)
myAttSample <- size(myAttSample, c = 0)
myAttSample$n
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.5")
myAttSample = att_sample(alpha=0.05, popn=12000000, popdev=120000).size(c=0)
myAttSample.n
```
<!-- ADA:END -->
To build a margin for one error, we may increase the critical region to $\{k | k \leq 1\}$, resulting in a sample size of $n =$ 473.

```{r}
myAttSample <- size(myAttSample, c = 1)
myAttSample$n
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.5")
myAttSample.size(c=1).n
```
<!-- ADA:END -->

In MUS, we increase the critical region by anticipating on the expected error (in monetary terms) in the population. Thus, in applications where selected items are either completely correct or completely incorrect, it is easier to calculate the minimum required sample size using the `att_obj` than using the `mus_obj`.

### Exercise 4.6. Case: Accounts receivable circularization

We start by setting up the MUS object `ar` (for Accounts Receivable) in `R` with the `FSaudit` package and first verify that the number of sampling units `popn` and the total book value `popBv` of the sampling frame match those in the population. Notice that these statistics are calculated as soon as the object is loaded with the detail amounts.

<!-- ADA:BEGIN lang=python mode=override kind=narrative -->
We start by setting up the MUS object `ar` (for Accounts Receivable) and first
verify that the number of sampling units `popn` and the total book value
`popBv` of the sampling frame match those in the population. Notice that these
statistics are calculated as soon as the object is loaded with the detail
amounts.
<!-- ADA:END -->

<!-- ADA:REQUIRES capability=fsaudit -->
```{r}
ar <- mus_obj(
  bv = accounts_receivable$amount,
  id = accounts_receivable$invoice
)
ar$popn
ar$popBv
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.6")
accounts_receivable = load_dataset("accounts_receivable")
ar = mus_sample(
  bv=accounts_receivable["amount"],
  id=accounts_receivable["invoice"],
)
ar.popn, ar.popBv
```
<!-- ADA:END -->
Sample size calculation is invoked with the relevant values from the *Case: Accounts receivable circularization*.

```{r}
ar <- size(ar,
  cl = 0.95, pm = 450000, ee = 100000,
  evalMeth = "Stringer"
)
ar$n
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.6")
ar.size(cl=0.95, pm=450000, ee=100000, evalMeth="Stringer").n
```
<!-- ADA:END -->
Compare this with the sample size calculated using fixed-attribute sampling.

```{r}
ar2 <- att_obj(alpha = 0.05, popn = 13500000, popdev = 450000)
ar2 <- size(ar2, c = 1)
ar2$n

ar2 <- size(ar2, c = 2)
ar2$n
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.6")
ar2 = att_sample(alpha=0.05, popn=13500000, popdev=450000).size(c=1)
ar2.n
ar2.size(c=2).n
```
<!-- ADA:END -->

We can therefore infer that with an expected error of `ee = 100000`, we can tolerate between one and two 100% errors.

### Exercise 4.7. Multiple hits with random selection

If we draw a sample of size $n = 1,100$ from the `accounts_receivable` population, the largest sampling units have an almost 100% inclusion probability and with a selection method such as `random` are likely to be hit more than once.

For example, the data file `accounts_receivable`, that comes with the `FSaudit` package, has the following six largest amounts:

<!-- ADA:REQUIRES capability=fsaudit -->
```{r}
accounts_receivable[order(-accounts_receivable$amount), ][1:6, 1:3]
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.7")
accounts_receivable.sort_values("amount", ascending=False).iloc[:6, :3]
```
<!-- ADA:END -->

We set up a new `mus_obj` and use a materiality of `pm = 36730` to arrive at a sample size of 1100.

```{r}
multiple <- mus_obj(
  bv = accounts_receivable$amount,
  id = accounts_receivable$invoice,
  pm = 36730
)
multiple <- size(multiple)
multiple$n
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.7")
multiple = mus_sample(
  bv=accounts_receivable["amount"],
  id=accounts_receivable["invoice"],
  pm=36730,
).size()
multiple.n
```
<!-- ADA:END -->

We select the sample with the selection method `selMeth = "random"` and order it in decreasing order, displaying the six largest book values.

```{r}
multiple <- select(multiple, selMeth = "random", seed = 1)
sample <- multiple$sample
head(sample[order(-sample$bv), ])
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.7")
multiple.select(selMeth="random", seed=1)
sample = multiple.sample
sample.sort_values("bv", ascending=False).head()
```
<!-- ADA:END -->
This example demonstrates that invoices 201719763 and 201710344 were selected twice.

### Exercise 4.8. Stringer bound

We continue the case study sample, and select the sample, now with the "randomized fixed" selection method.

<!-- ADA:REQUIRES capability=fsaudit -->
```{r}
ar <- select(ar,
  selMeth = "randomized.fixed",
  seed = 345
)
head(ar$sample)
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.8")
ar.select(selMeth="randomized.fixed", seed=345)
ar.sample.head()
```
<!-- ADA:END -->

Before we can evaluate the sample, we must first provide audit values. These should be provided in a list of the same order as the list of sample book values. We first copy the list of book values into a data frame.

```{r}
myResults <- data.frame(item = ar$sample$item, av = ar$sample$bv)
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.8")
myResults = ar.sample.loc[:, ["item", "bv"]].rename(columns={"bv": "av"}).copy()
myResults
```
<!-- ADA:END -->
Table 5.4. lists the three invoices (items 16, 52, and 124), that are assumed to be in error.
```{r}
myResults[c(16, 52, 124), ]
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.8")
myResults.iloc[[15, 51, 123], :]
```
<!-- ADA:END -->
We update the audit values of the erroneous items.
```{r}
myResults[16, 2] <- 4438.82
myResults[52, 2] <- 0
myResults[124, 2] <- 5531.38
myResults[c(16, 52, 124), ]
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.8")
myResults.iloc[15, 1] = 4438.82
myResults.iloc[51, 1] = 0
myResults.iloc[123, 1] = 5531.38
myResults.iloc[[15, 51, 123], :]
```
<!-- ADA:END -->

The updated list of audit values is then submitted to the MUS object for evaluation.

```{r}
ar <- evaluate(ar, av = myResults$av, evalMeth = "stringer")
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.8")
ar.evaluate(av=myResults["av"], evalMeth="stringer")
```
<!-- ADA:END -->

The results of the Stringer bound evaluation as presented in Tables 5.6 and 5.7 are stored in the `Precision calculation` attribute.

```{r}
ar$evalResults$Over$`Precision calculation`
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.8")
ar.eval_results["Over"]["Precision calculation"]
```
<!-- ADA:END -->

The upper bounds $M_U$ in Table 5.5 are calculated with the `upper` function from the `FSaudit` package.

```{r}
upper(k = 0, popn = 13500000, n = 145, alpha = .05)
upper(k = 1, popn = 13500000, n = 145, alpha = .05)
upper(k = 2, popn = 13500000, n = 145, alpha = .05)
upper(k = 3, popn = 13500000, n = 145, alpha = .05)
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
upper_bound(k=0, popn=13500000, n=145, alpha=0.05)
upper_bound(k=1, popn=13500000, n=145, alpha=0.05)
upper_bound(k=2, popn=13500000, n=145, alpha=0.05)
upper_bound(k=3, popn=13500000, n=145, alpha=0.05)
```
<!-- ADA:END -->

```{r hyp-worked-calculation-values-mus-sample-size, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
if (!requireNamespace("jsonlite", quietly = TRUE)) {
  stop("jsonlite is required to emit manuscript calculation values.")
}

find_repo_root <- function(start = getwd()) {
  current <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(current, "pyproject.toml")) || file.exists(file.path(current, ".git"))) {
      return(current)
    }
    parent <- dirname(current)
    if (identical(parent, current)) {
      stop("Unable to locate repository root from working directory: ", start)
    }
    current <- parent
  }
}

repo_root <- find_repo_root()
values_output_dir <- file.path(
  repo_root,
  "generated",
  "cache",
  "manuscript-calculation-values"
)
dir.create(values_output_dir, recursive = TRUE, showWarnings = FALSE)

write_values_entry <- function(file_name, entry) {
  values_output_path <- file.path(values_output_dir, file_name)
  jsonlite::write_json(entry, path = values_output_path, auto_unbox = TRUE, pretty = TRUE)
  message("Wrote manuscript calculation values: ", values_output_path)
}

X_total <- 13500000
expected_error <- 100000
performance_materiality <- 450000

att_k1 <- att_obj(alpha = 0.05, popn = X_total, popdev = performance_materiality)
att_k1 <- size(att_k1, c = 1)
n_1 <- att_k1$n

att_k0 <- att_obj(alpha = 0.05, popn = X_total, popdev = performance_materiality)
att_k0 <- size(att_k0, c = 0)
n_0 <- att_k0$n

att_k2 <- att_obj(alpha = 0.05, popn = X_total, popdev = performance_materiality)
att_k2 <- size(att_k2, c = 2)
n_2 <- att_k2$n

att_k3 <- att_obj(alpha = 0.05, popn = X_total, popdev = performance_materiality)
att_k3 <- size(att_k3, c = 3)
n_3 <- att_k3$n

sampling_interval_k1 <- X_total / n_1
sampling_interval_k2 <- X_total / n_2
expected_error_k2 <- 2 * sampling_interval_k2

sampling_interval_k0 <- X_total / n_0
sampling_interval_k3 <- X_total / n_3
expected_error_k0 <- 0
expected_error_k1 <- sampling_interval_k1
expected_error_k3 <- 3 * sampling_interval_k3

mu_0 <- upper(k = 0, popn = X_total, n = ar$n, alpha = .05)

mu_1 <- upper(k = 1, popn = X_total, n = ar$n, alpha = .05)
mu_2 <- upper(k = 2, popn = X_total, n = ar$n, alpha = .05)
mu_3 <- upper(k = 3, popn = X_total, n = ar$n, alpha = .05)

cell_t1 <- 1.00
cell_t2 <- 0.20
cell_t3 <- 0.04
cell_tbar_1 <- cell_t1
cell_tbar_2 <- (cell_t1 + cell_t2) / 2
cell_tbar_3 <- (cell_t1 + cell_t2 + cell_t3) / 3
cell_si <- X_total / ar$n

cell_simple_1 <- mu_1 * cell_tbar_1
cell_load_1 <- mu_0 + cell_t1 * cell_si
cell_stage_1 <- max(cell_simple_1, cell_load_1)

cell_simple_2 <- mu_2 * cell_tbar_2
cell_load_2 <- cell_stage_1 + cell_t2 * cell_si
cell_stage_2 <- max(cell_simple_2, cell_load_2)

cell_simple_3 <- mu_3 * cell_tbar_3
cell_load_3 <- cell_stage_2 + cell_t3 * cell_si
cell_stage_3 <- max(cell_simple_3, cell_load_3)

cell_projected_misstatement <- cell_si * (cell_t1 + cell_t2 + cell_t3)

k_value <- 1
n_star <- (k_value - n_1 / (n_2 - n_1)) / ((expected_error / X_total) - 1 / (n_2 - n_1))
n_required <- ceiling(n_star)
t_tol <- k_value + (performance_materiality - mu_1) / (mu_2 - mu_1)
partial_tainting_pct <- (t_tol - k_value) * 100

write_values_entry(
  "hyp.mus.sample_size.json",
  list(
    id = "hyp.mus.sample_size",
    kind = "worked_calculation",
    source_notebook = "notebooks/support/hypothesis-testing/support.Rmd",
    source_context = "Exercise 4.6 and Exercise 4.8 MUS sample-size interpolation and tolerable tainting",
    source_dataset = "FSaudit::accounts_receivable",
    target_snippet = "generated/worked-calculations/hyp-mus-sample-size.tex",
    tolerance = 1e-8,
    language_scope = "r",
    equation_labels = c("eq:MUS_sample_size"),
    values = list(
      list(id = "hyp.mus.population_book_total", role = "population_book_total", raw = X_total, format = "integer"),
      list(id = "hyp.mus.performance_materiality", role = "performance_materiality", raw = performance_materiality, format = "integer"),
      list(id = "hyp.mus.expected_error", role = "expected_error", raw = expected_error, format = "integer"),
      list(id = "hyp.mus.n_1", role = "n_1", raw = n_1, format = "integer"),
      list(id = "hyp.mus.n_2", role = "n_2", raw = n_2, format = "integer"),
      list(id = "hyp.mus.sampling_interval_k1", role = "sampling_interval_k1", raw = sampling_interval_k1, format = "number:2"),
      list(id = "hyp.mus.sampling_interval_k2", role = "sampling_interval_k2", raw = sampling_interval_k2, format = "number:2"),
      list(id = "hyp.mus.expected_error_k2", role = "expected_error_k2", raw = expected_error_k2, format = "number:2"),
      list(id = "hyp.mus.mu_1", role = "mu_1", raw = mu_1, format = "integer"),
      list(id = "hyp.mus.mu_2", role = "mu_2", raw = mu_2, format = "integer"),
      list(id = "hyp.mus.k_value", role = "k_value", raw = k_value, format = "integer"),
      list(id = "hyp.mus.n_star", role = "n_star", raw = n_star, format = "number:1"),
      list(id = "hyp.mus.n_required", role = "n_required", raw = n_required, format = "integer"),
      list(id = "hyp.mus.t_tol", role = "t_tol", raw = t_tol, format = "number:5"),
      list(id = "hyp.mus.partial_tainting_pct", role = "partial_tainting_pct", raw = partial_tainting_pct / 100, format = "percentage:3")
    )
  )
)

write_values_entry(
  "hyp.mus.attribute_sample_sizes_rows.json",
  list(
    id = "hyp.mus.attribute_sample_sizes_rows",
    kind = "worked_calculation",
    source_notebook = "notebooks/support/hypothesis-testing/support.Rmd",
    source_context = "Sample-size table rows for MUS interpolation example",
    source_dataset = "FSaudit::accounts_receivable",
    target_snippet = "generated/worked-calculations/hyp-mus-attribute-sample-sizes-rows.tex",
    tolerance = 1e-8,
    language_scope = "r",
    values = list(
      list(id = "hyp.mus.n_0", role = "n_0", raw = n_0, format = "integer"),
      list(id = "hyp.mus.n_1", role = "n_1", raw = n_1, format = "integer"),
      list(id = "hyp.mus.n_2", role = "n_2", raw = n_2, format = "integer"),
      list(id = "hyp.mus.n_3", role = "n_3", raw = n_3, format = "integer"),
      list(id = "hyp.mus.sampling_interval_k0", role = "sampling_interval_k0", raw = sampling_interval_k0, format = "number:2"),
      list(id = "hyp.mus.sampling_interval_k1", role = "sampling_interval_k1", raw = sampling_interval_k1, format = "number:2"),
      list(id = "hyp.mus.sampling_interval_k2", role = "sampling_interval_k2", raw = sampling_interval_k2, format = "number:2"),
      list(id = "hyp.mus.sampling_interval_k3", role = "sampling_interval_k3", raw = sampling_interval_k3, format = "number:2"),
      list(id = "hyp.mus.expected_error_k0", role = "expected_error_k0", raw = expected_error_k0, format = "number:2"),
      list(id = "hyp.mus.expected_error_k1", role = "expected_error_k1", raw = expected_error_k1, format = "number:2"),
      list(id = "hyp.mus.expected_error_k2", role = "expected_error_k2", raw = expected_error_k2, format = "number:2"),
      list(id = "hyp.mus.expected_error_k3", role = "expected_error_k3", raw = expected_error_k3, format = "number:2")
    )
  )
)

write_values_entry(
  "hyp.cell.evaluation_steps.json",
  list(
    id = "hyp.cell.evaluation_steps",
    kind = "worked_calculation",
    source_notebook = "notebooks/support/hypothesis-testing/support.Rmd",
    source_context = "Cell-bounds stepwise comparison and resulting upper bound",
    source_dataset = "FSaudit::accounts_receivable",
    target_snippet = "generated/worked-calculations/hyp-cell-evaluation-steps.tex",
    tolerance = 1e-8,
    language_scope = "r",
    values = list(
      list(id = "hyp.cell.mu_0", role = "mu_0", raw = mu_0, format = "integer"),
      list(id = "hyp.cell.mu_1", role = "mu_1", raw = mu_1, format = "integer"),
      list(id = "hyp.cell.mu_2", role = "mu_2", raw = mu_2, format = "integer"),
      list(id = "hyp.cell.mu_3", role = "mu_3", raw = mu_3, format = "integer"),
      list(id = "hyp.cell.t_1", role = "t_1", raw = cell_t1, format = "number:2"),
      list(id = "hyp.cell.t_2", role = "t_2", raw = cell_t2, format = "number:2"),
      list(id = "hyp.cell.t_3", role = "t_3", raw = cell_t3, format = "number:2"),
      list(id = "hyp.cell.tbar_1", role = "tbar_1", raw = cell_tbar_1, format = "number:2"),
      list(id = "hyp.cell.tbar_2", role = "tbar_2", raw = cell_tbar_2, format = "number:2"),
      list(id = "hyp.cell.tbar_3", role = "tbar_3", raw = cell_tbar_3, format = "number:4"),
      list(id = "hyp.cell.sampling_interval", role = "sampling_interval", raw = cell_si, format = "number:2"),
      list(id = "hyp.cell.simple_1", role = "simple_1", raw = cell_simple_1, format = "number:2"),
      list(id = "hyp.cell.load_1", role = "load_1", raw = cell_load_1, format = "number:2"),
      list(id = "hyp.cell.stage_1", role = "stage_1", raw = cell_stage_1, format = "number:2"),
      list(id = "hyp.cell.simple_2", role = "simple_2", raw = cell_simple_2, format = "number:2"),
      list(id = "hyp.cell.load_2", role = "load_2", raw = cell_load_2, format = "number:2"),
      list(id = "hyp.cell.stage_2", role = "stage_2", raw = cell_stage_2, format = "number:2"),
      list(id = "hyp.cell.simple_3", role = "simple_3", raw = cell_simple_3, format = "number:2"),
      list(id = "hyp.cell.load_3", role = "load_3", raw = cell_load_3, format = "number:2"),
      list(id = "hyp.cell.stage_3", role = "stage_3", raw = cell_stage_3, format = "number:2"),
      list(id = "hyp.cell.projected_misstatement", role = "projected_misstatement", raw = cell_projected_misstatement, format = "number:2")
    )
  )
)
```

### Exercise 4.9. Cell evaluation

The results of the cell evaluation method, as presented in Table 5.8, can also be obtained, by changing the evaluation method.

<!-- ADA:REQUIRES capability=fsaudit -->
```{r}
options(width = 70)
ar <- evaluate(ar, av = myResults$av, evalMeth = "cell")
ar$evalResults$Over$`Precision calculation`
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.9")
pd.set_option("display.width", 70)
ar.evaluate(av=myResults["av"], evalMeth="cell")
ar.eval_results["Over"]["Precision calculation"]
```
<!-- ADA:END -->

### Exercise 4.10. PPS estimation

Finally, we present results from the `pps` evaluation method. For this purpose, we use audit values stored in the variable `av2`. The total error amount reflected in `av2` is 450,000.

<!-- ADA:REQUIRES capability=fsaudit -->
```{r}
ar <- evaluate(ar,
  av = myResults$av2,
  evalMeth = "pps"
)
ar$evalResults$`Error estimate`
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.10")
myResults["av2"] = accounts_receivable.set_index("invoice").loc[myResults["item"], "av2"].to_numpy()
ar.evaluate(av=myResults["av2"], evalMeth="pps")
ar.eval_results["Error estimate"]
```
<!-- ADA:END -->
A two-sided prediction interval around the PPS estimate is calculated according to Equations 5.6 and 5.7.

```{r}
ar$evalResults$`pps estimate`
ar$evalResults$`Lower bound`
ar$evalResults$`Upper bound`
```
<!-- ADA:BEGIN lang=python mode=override kind=code -->
```{r}
ada_set_context("4.10")
ar.eval_results["pps estimate"]
ar.eval_results["Lower bound"]
ar.eval_results["Upper bound"]
```
<!-- ADA:END -->



<!-- SUPPORT-ONLY:START -->

## Private book-support calculations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(pillar.sigfig = 4)
library(FSaudit)
library(ggplot2)
if (requireNamespace("kthemes", quietly = TRUE)) {
  library(kthemes)
  theme_set(theme_kpmg())
} else {
  theme_set(theme_grey())
}
```

## Attribute Sampling
*Case: Information provided by the entity*.
```{r}
mySample <- att_obj(alpha = .1,
                    popdev = 60,
                    popn = 1200,
                    c = 0)
```

```{r}
mySample <- size(mySample)
mySample$n
```
```{r}
mySample <- size(mySample, c = 2)
mySample$n
```
```{r}
phyper(q = 1, m = 60, n = 1140, k = 45)
phyper(q = 2, m = 60, n = 1140, k = 102)
```

```{r}
phyper(0, m = 24, n = 1176, k = 45, lower.tail = FALSE)
```

```{r}
k <- seq(0, 59, by = 1)

y <- phyper(0, m = k, n = 1200 - k, k = 45, lower.tail = FALSE)
s2 <- "Type II error"
df <- data.frame(number_errors = k, prob_accept = y, series = s2)

occ <-
  ggplot(df) +
  geom_line(aes(x = number_errors, y = prob_accept, colour = series), 
            show.legend = FALSE, size = 1) +
  # theme(axis.text.y=element_blank()) +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12)) +
  theme(legend.title=element_blank()) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.20)) +
  scale_x_continuous(breaks = seq(0, 70, by = 10)) +
  xlab("Number of errors in population") +
  ylab("Probability of rejection")
occ
save_book_figure(occ, "TypeIIerror.pdf")
```
OCC for three fixed attribute plans

```{r}
# create OCC graph for discovery, fixed n = 77 and fixed n = 45

p <- seq(0, 0.07, by = 0.0001)

y2 <- dbinom(0, 77, p) + dbinom(1, 77, p)
s2 <- "fixed, n = 77"
y1 <- dbinom(0, 45, p)
s1 <- "discovery, n = 45"
y3 <- dbinom(0, 105, p) + dbinom(1, 105, p) + dbinom(2, 105, p)
s3 <- "fixed, n = 105"

df1 <- data.frame(error_rate = p, prob_accept = y1, series = s1)
df2 <- data.frame(error_rate = p, prob_accept = y2, series = s2)
df3 <- data.frame(error_rate = p, prob_accept = y3, series = s3)

dftot <- rbind(df1, df2, df3)

# To make sure that labels within the legend are not in alphabetical order.
dftot$series <- factor(dftot$series, levels = c(s1, s2, s3))

occ <-
  ggplot(dftot) +
  geom_line(aes(x = error_rate, y = prob_accept, colour = series), size=1) +
  # theme(axis.text.y=element_blank()) +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12)) +
  theme(legend.title=element_blank()) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.10)) +
  scale_x_continuous(breaks = seq(0, 0.1, by = 0.01)) +
  xlab("Population deviation rate") +
  ylab("Probability of rejecting the null") +
  geom_vline(xintercept = 0.05)
occ
save_book_figure(occ, "three_attribute_occ.pdf")
```

OCC for double vs discovery and fixed

```{r}
# create OCC graph for discovery, fixed n = 77 and fixed n = 45

p_accept <- function(n, p) {
  # chance of finding k=0
  pk0 <- dbinom(x = 0, size = n, prob = p)
  # P(k=1)
  pk1 <- dbinom(x = 1, size = n, prob = p)
  # chance of finding first k=1, then in second sample k=0
  pk10 <- pk1 * pk0
  # Total probability to accept
  p_accept <- pk0 + pk10
  return(p_accept)
}

p <- seq(0, 0.07, by = 0.0001)

y3 <- dbinom(0, 77, p) + dbinom(1, 77, p)
s3 <- "fixed, n = 77"
y1 <- dbinom(0, 45, p)
s1 <- "discovery, n = 45"
y2 <- p_accept(49, p)
s2 <- "double, n0 = n1 = 49"

df1 <- data.frame(error_rate = p, prob_accept = y1, series = s1)
df2 <- data.frame(error_rate = p, prob_accept = y2, series = s2)
df3 <- data.frame(error_rate = p, prob_accept = y3, series = s3)

dftot <- rbind(df1, df2, df3)

# To make sure that labels within the legend are not in alphabetical order.
dftot$series <- factor(dftot$series, levels = c(s1, s2, s3))

occ <-
  ggplot(dftot) +
  geom_line(aes(x = error_rate, y = prob_accept, colour = series), size=1) +
  # theme(axis.text.y=element_blank()) +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12)) +
  theme(legend.title=element_blank()) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.10)) +
  scale_x_continuous(breaks = seq(0, 0.1, by = 0.01)) +
  xlab("Population deviation rate") +
  ylab("Probability of rejecting the null") +
  geom_vline(xintercept = 0.05)
occ
save_book_figure(occ, "att_double_occ.pdf")
```


Table 5.1
```{r}
alpha <- 0.1
n <- 102
popn <- 1200
k <- c(0:3)
kdivn <- k / n
Nkdivn <- popn * k / n
M_U <- c(0, 0, 0, 0)
for (i in 1:4) {M_U[i] <- upper(i - 1, popn, n, alpha)}
p_U <- M_U / popn
(df <- data.frame(k, kdivn, Nkdivn, M_U, p_U))
```






The examples above show how samples sizes can be calculated for various critical regions. The default distribution to be used is the hypergeometric, but calculations can also be performed with the approximating binomial and Poisson distributions. In that case, the distribution needs to be specifically entered as an argument, and the function does not require a value for $popn$. 
```{r}
mySample2 <- att_obj(alpha = .1,
                tdr = .05,
                c = 2,
                dist = "binom")
mySample2 <- size(mySample2)
mySample2$n
```
Same for the Poisson distribution
```{r}
mySample3 <- att_obj(alpha = .1,
                tdr = .05,
                c = 2,
                dist = "pois")
mySample3 <- size(mySample3)
mySample3$n
```
Note that the approximating binomial and Poisson distributions lead to sample sizes that are unnecessarily larger than those calculated with the hypergeometric distribution.

We can evaluate an attribute sample with the $p$-value in the following way. We first calculate $P(k ≤ 1,n = 45, N = 1200, M = 60)$
```{r}
phyper(q = 1, m = 60, n = 1140, k = 45)
```
Similarly, when we designed a plan that tolerates two errors and
we find two, the $p$-value $P(k ≤ 2, n = 102, N = 1200, M = 60)$ equals:
```{r}
phyper(q = 2, m = 60, n = 1140, k = 102)
```

## Monetary Unit Sampling

### Case: European innovation subsidies
The sample size in the *Case: European innovation subsidies* depends on the critical region chosen. When the null hypothesis $H_0 : M \geq 120,000$ is rejected when the sample yields no errors, the critical region is $\{k | k = 0\}$. We first create an `mus_obj` object, and load it with the parameters.
```{r}
subsidies <- mus_obj(cl = 0.95, 
                     popBv = 12000000, 
                     pm = 120000)
```
Now we calculate the minimum required sample size as follows:
```{r}
subsidies <- size(subsidies,
                  ee = 0)
subsidies$n
```
```{r}
myAttSample <- att_obj(alpha = 0.05, popn = 12000000, popdev = 120000)
myAttSample <- size(myAttSample, c = 0)
myAttSample$n
```
```{r}
myAttSample <- size(myAttSample, c = 1)
myAttSample$n
```

# Accounts receivable circularization
For the sample on accounts receivable in the *Case: Accounts receivable circularization*, we again start with assigning the `mus_obj` object type to the object `ar`.
```{r}
ar <- mus_obj(bv = accounts_receivable$amount,
              id = accounts_receivable$invoice)
```
In this first step we have assigned the unique `id` of the sampling units in the sampling frame, the `invoice' field, and the book values `bv' of the sampling units. As soon as these arguments are assigned, `R' calculates the number of sampling units in the sampling frame `popn', and the total book value `popBv', and stores these values in the `ar` object for future purposes.
```{r}
ar$popn
ar$popBv
```

```{r}
ar <- size(ar, cl = 0.95, pm = 450000, ee = 100000)
ar$n

ar2 <- att_obj(alpha = 0.05, popn = 13500000, popdev = 450000)
ar2 <- size(ar2, c = 1)
ar2$n

ar2 <- size(ar2, c = 2)
ar2$n
```
Selection methods with a probability proportional to size are implemented in the `FSaudit` package with the `select` function. The default method is the randomized fixed interval selection method:
```{r}
ar <- select(ar,
             selMeth = "randomized.fixed",
             seed = 345)
head(ar$sample)
```



To evaluate the selected and audited sampling units, we first obtain their audit values in a data frame. 

```{r}
myResults <- 
  accounts_receivable[match(ar$sample$item, 
                            accounts_receivable$invoice), 
                      c("invoice", "av2")]
options(pillar.sigfig = 6)
head(myResults, 6)
```
Once we have audit values for each selected book value, we can evaluate the sample.
```{r}
options(digits = 8)
ar <- evaluate(ar,
               av = myResults$av2,
               evalMeth = "cell")
```
The results of the evaluation are stored in the \ttblue{ar} object and can be obtained as follows:
```{r}
ar$evalResults$Over$`Projected misstatement`
ar$evalResults$Over$`Sample size used`
ar$evalResults$Over$`Nonzero differences`
ar$evalResults$Over$`Cell width`
ar$evalResults$Over$`Basic precision`
ar$evalResults$Over$`Precision achieved`
ar$evalResults$Over$`Upper confidence bound`
```
The precision calculations have been summarized in a separate data frame:
```{r}
ar$evalResults$Over$`Precision calculation`

```
If you are curious which items contain the differences, you can obtain their `id` and `bv` as follows. We first get a vector of the sampling units that have nonzero differences. Notice that `R` prints the results immediately because we put the assignment of the `errors` variable in brackets.
```{r}
(errors <- which(ar$sample$bv != ar$av))
```
This vector of integers can then be used to look up the unique id, book value and audit value.
```{r}

misstatements <- data.frame(id = ar$sample$item[errors], 
                            bv = ar$sample$bv[errors], 
                            av = ar$av[errors])
misstatements$diff  <- misstatements$bv - misstatements$av 
misstatements$taint <- misstatements$diff / misstatements$bv
misstatements
```




Use Stringer bound evaluation

```{r}
ar <- evaluate(ar,
               av = myResults$av2,
               evalMeth = "stringer")
```
Errors found
```{r}
ar$evalResults$Over$`Sampling interval`
ar$evalResults$Over$`Precision achieved`
ar$evalResults$Over$`Upper confidence bound`
```
Calculations
```{r}
ar$evalResults$Over$`Precision calculation`
```
We conclude with the results of the `pps` evaluation method.

```{r}
ar <- evaluate(ar,
               av = myResults$av2,
               evalMeth = "pps")
```
Errors found
```{r}
ar$evalResults$`Nonzero diff`
ar$evalResults$`pps estimate`
ar$evalResults$`Error estimate`
ar$evalResults$Precision
ar$evalResults$`Effective df`
ar$evalResults$`Lower bound`
ar$evalResults$`Upper bound`

```
<!-- SUPPORT-ONLY:END -->
