## Exercise 5.26. The Cochrane--Orcutt method

A practical and widely used fix is the `Cochrane-Orcutt method`, designed specifically to correct for first-order autocorrelation (AR(1)). This method estimates the autocorrelation parameter and transforms the regression model by subtracting a multiple of lagged values from both the dependent and independent variables. The transformed model can then be estimated by ordinary least squares, restoring valid inference under autocorrelation.

We first estimate an AR(1) model for the residuals from `mod.3`.

```{r}
(rho <- arima(ussteamco.mod.3$residuals,
              order = c(1, 0, 0))$coef[1])
```

Then we transform all variables using the estimated $\rho$. For each interaction, we first construct the interaction from the original variables at times $t$ and $t-1$, and then apply the AR(1) transformation to that interaction:

$$x_t s_t - \rho x_{t-1}s_{t-1}.$$

```{r}
n <- nrow(USSteamCoEstim2)
trans <- USSteamCoEstim2[-1, ]  # t
lag   <- USSteamCoEstim2[-n, ]  # t-1

trans$revenue_adj <- trans$revenue - rho * lag$revenue

trans$production_adj <- trans$production - rho * lag$production
trans$heatDD_adj     <- trans$heatDD     - rho * lag$heatDD
trans$coolDD_adj     <- trans$coolDD     - rho * lag$coolDD
trans$summer_adj     <- trans$summer     - rho * lag$summer

trans$prod_summer_adj <- trans$production * trans$summer -
  rho * lag$production * lag$summer
trans$heat_summer_adj <- trans$heatDD * trans$summer -
  rho * lag$heatDD * lag$summer
trans$cool_summer_adj <- trans$coolDD * trans$summer -
  rho * lag$coolDD * lag$summer
```

We fit a new model using the Cochrane--Orcutt method.

```{r}
ussteamco.mod.4 <- lm(revenue_adj ~
                        production_adj + heatDD_adj + coolDD_adj +
                        summer_adj + prod_summer_adj +
                        heat_summer_adj + cool_summer_adj,
                      data = trans)

summary(ussteamco.mod.4)
```

This manual approach follows the Cochrane--Orcutt transformation:

$$y_t - \rho y_{t-1} = \beta_0(1 - \rho) + \sum_j \beta_j(x_{jt} - \rho x_{j,t-1}) + u_t.$$

It discards the first observation, which cannot be transformed because no lagged observation is available. After fitting the transformed model, we check whether autocorrelation is reduced by inspecting the `pacf` plot and running the Breusch--Godfrey test.

```{r}
pacf(ussteamco.mod.4$residuals)
bgtest(ussteamco.mod.4, order = 3)
```
